---
title: "Data Source Investigation"
author: "Grant Foster"
date: "2024-03-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document is for investigating network data structure (to inform methods), as well as to create a conceptual figure.

```{r}
library(tidyverse)
library(magrittr)
library(igraph)
library(BIEN)
library(PVR)
library(ape)
library("phytools")
library(picante)
```


```{r}
dat <- read.csv("Data.nosync/ATLANTIC_frugivory.csv")
dat %<>% dplyr::filter(., Frugivore_Species != "Carollia castanea") #This bat has an incorrect gape size, so it's filtered out
dat <- dplyr::select(dat, -ID, -Latitude, -Longitude, -Study_Location, -Precision, -Study_Method, -Study.reference, -Doi.Link, -Frug_Population_Trend, -Frug_Migration_status)
dat %<>% unique(.)
birds <- dplyr::filter(dat, Frug_Class=="Aves")
```

Number of unique species and interactions
```{r}
length(unique(dat$Frugivore_Species))
length(unique(dat$Plant_Species))

length(unique(birds$Frugivore_Species))
length(unique(birds$Plant_Species))
```

Degree disributions

```{r}
hist(table(birds$Frugivore_Species))
table(table(birds$Frugivore_Species)==1)
median(table(dat$Frugivore_Species))
```


```{r}
hist(table(birds$Plant_Species))
range(table(birds$Plant_Species))
table(table(birds$Plant_Species)==1)
median(table(dat$Plant_Species))
```


```{r}
load("BIEN_subtree.Rda")
plant_litter <- PVR::PVRdecomp(phy=BIEN_subtree, type="newick", scale=TRUE)

sum((round((plant_litter@Eigen$values)/sum((plant_litter@Eigen$values)),3)*100)[1:4]) #first three vectors contain about 42.6% of variation :(

plant_PhyEig <-data.frame(BIEN_subtree$tip.label, plant_litter@Eigen$vectors[,1:5])
colnames(plant_PhyEig) <- c("Plant_Species", paste(colnames(plant_PhyEig)[2:ncol(plant_PhyEig)], "PlDecomp", sep=""))

plant_PhyEig$Plant_Species <- gsub(pattern="_", replace=" ", x=plant_PhyEig$Plant_Species)

dat <- left_join(dat, plant_PhyEig, by="Plant_Species")
birds <- dplyr::filter(dat, Frug_Class=="Aves") %>% #Filter to birds
  dplyr::filter(., grepl(" ", Frugivore_Species)==TRUE) #Filter out any bird records only to the genus level
```

```{r}

dmatrix <- cophenetic(BIEN_subtree)
dmatrix <- dmatrix/max(dmatrix)

birds$plant_Tips <- gsub(pattern=" ", replace="_", x=birds$Plant_Species)

birds_phylo<- dplyr::filter(birds, plant_Tips %in% BIEN_subtree$tip.label)

comm <- as.data.frame.matrix(table(birds_phylo$Frugivore_Species, birds_phylo$plant_Tips))


test_mpd <- picante::ses.mpd(comm, dmatrix, null.model = "independentswap", abundance.weighted=FALSE, runs = 999, iterations = 1000) 

phylospec <- dplyr::filter(test_mpd, ntaxa > 1)
dplyr::filter(phylospec, mpd.obs.p<0.05) #Number of species more significantly phylogenetically specialized than a random draw. 

```

Filling in missing data points with AvoNet. 
Turns out, AvoNet doesn't actually provide gape size. It does have a number of other relate beak size measures, but without some assumptions I can't use those to fill it in. 
```{r}
bm <- birds %>% dplyr::select(., Frugivore_Species, Frug_Body_Mass) %>% unique() %>% dplyr::filter(., is.na(Frug_Body_Mass)==TRUE) %>% dplyr::filter(., grepl(" ", Frugivore_Species)==TRUE)


gapes <- birds %>% dplyr::select(., Frugivore_Species, Frug_Mean_Gape_Size) %>% unique() %>% dplyr::filter(., is.na(Frug_Mean_Gape_Size)==TRUE) %>% dplyr::filter(., grepl(" ", Frugivore_Species)==TRUE)
```


```{r}
birdsp <- birds %>% dplyr::select(., Frugivore_Species, Frug_Mean_Gape_Size, Frug_Body_Mass) %>% unique() %>% dplyr::filter(., grepl(" ", Frugivore_Species)==TRUE)
Avonet <- read.csv(file="Data.nosync/AvoNet/Avonet1_BirdLife.csv")

Avonet <- Avonet %>% dplyr::filter(., Species1 %in% birdsp$Frugivore_Species) %>% dplyr::select(., Species1, Beak.Length_Nares, Beak.Width, Beak.Depth)
colnames(Avonet) <- c("Frugivore_Species", "Beak_len", "Beak_wid", "Beak_dep")

birdsp <- left_join(birdsp, Avonet, by="Frugivore_Species")

DL <- lm(formula=Frug_Mean_Gape_Size~Beak_len+Beak_dep, data=birdsp)
L <- lm(formula=Frug_Mean_Gape_Size~Beak_len, data=birdsp)
D <- lm(formula=Frug_Mean_Gape_Size~Beak_dep, data=birdsp)
DL_int <- lm(formula=Frug_Mean_Gape_Size~Beak_len+Beak_dep+Beak_len*Beak_dep, data=birdsp)
DL_quad <- lm(formula=Frug_Mean_Gape_Size~exp(Beak_len)+Beak_len+Beak_dep, data=birdsp)
WidL <- lm(formula=Frug_Mean_Gape_Size~Beak_len+Beak_wid, data=birdsp)
WidLint <- lm(formula=Frug_Mean_Gape_Size~Beak_len+Beak_wid+Beak_len*Beak_wid, data=birdsp)


WidInt <- lm(formula=Frug_Mean_Gape_Size~Beak_len+Beak_dep+Beak_wid, data=birdsp)

WidInt2 <- lm(formula=Frug_Mean_Gape_Size~Beak_len+Beak_dep+Beak_wid+Beak_len*Beak_wid, data=birdsp)

extractAIC(L)
extractAIC(D)
extractAIC(DL) 
extractAIC(DL_int)
extractAIC(DL_quad)
extractAIC(WidL) 
extractAIC(WidLint)#Best AIC value
extractAIC(WidInt)
extractAIC(WidInt2)


birdsp$predGape <- predict(WidL, birdsp)

pdf(file="GapePredictions.pdf")
plot(x=birdsp$Frug_Mean_Gape_Size, y=birdsp$predGape, xlab="Measured Gape", ylab="Predicted Gape", xlim=c(0,40), ylim=c(0,40))
abline(a=0, b=1)
dev.off()

summary(WidL)

plot(WidL)

save(WidL, file="GapeModel.Rda")
plot(WidL)
```

Same thing for BIEN

```{r}
diam <- birds %>% dplyr::select(., Plant_Species, fruit_diameter) %>% unique() %>% dplyr::filter(., is.na(fruit_diameter)==TRUE) %>% dplyr::filter(., grepl(" ", Plant_Species)==TRUE)

len <- birds %>% dplyr::select(., Plant_Species, fruit_length) %>% unique() %>% dplyr::filter(., is.na(fruit_length)==TRUE) %>% dplyr::filter(., grepl(" ", Plant_Species)==TRUE)

birds$Fruit_color[birds$Fruit_color==""] <- NA
colr <- birds %>% dplyr::select(., Plant_Species, Fruit_color) %>% unique() %>% dplyr::filter(., is.na(Fruit_color)==TRUE) %>% dplyr::filter(., grepl(" ", Plant_Species)==TRUE)

plant_fills <- unique(c(diam$Plant_Species, len$Plant_Species, colr$Plant_Species))
library(BIEN)
```


```{r}
read.csv(file="Data.nosync/AvoNet/")
```


Making a conceptual figure
```{r}
set.seed(5)
fsp_ID <- sample(1:length(unique(birds$Frugivore_Species)), size=10, replace=FALSE)
frugs <- unique(birds$Frugivore_Species)

sNet <- dplyr::filter(birds, Frugivore_Species %in% frugs[fsp_ID]) %>% dplyr::filter(., Frugivore_Species %in% c("Saltator maximus", "Mimus saturninus")==FALSE)

length(unique(sNet$Plant_Species))
length(unique(sNet$Frugivore_Species))

table(sNet$Frugivore_Species)
```



```{r}
web <- data.frame(
    from = sNet$Frugivore_Species,
    to = sNet$Plant_Species
  ) %>% graph_from_data_frame(directed=FALSE) #Create a graph with all unique plant-frugivore edges

V(web)$type <- c(rep("Frug", 8), rep("Plant",42-8))
lyt <- layout.bipartite(web, types=c(rep(TRUE, 8), rep(FALSE,42-8)))



plot(web, layout=lyt)
```

