---
title: "Frugivory"
author: "Grant Foster"
date: "9/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(magrittr)
library(igraph)
library(pROC)
library(BIEN)
library(np)
library(PVR)
library(ape)
library(phytools)
library(tictoc)
```


```{r read in data}
tic()
dat <- read.csv("Data.nosync/ATLANTIC_frugivory.csv")
dat %<>% filter(., Frugivore_Species != "Carollia castanea") #This bat has an incorrect gape size
dat <- dplyr::select(dat, -ID, -Latitude, -Longitude, -Study_Location, -Precision, -Study_Method, -Study.reference, -Doi.Link, -Frug_Population_Trend, -Frug_Migration_status)

dat %<>% unique(.)
```

```{r set up igraph network eval=FALSE}

web <- data.frame(
    from = dat$Frugivore_Species,
    to = dat$Plant_Species
  ) %>% graph_from_data_frame(directed=FALSE) #Create a graph with all unique plant-frugivore edges

mat <-get.adjacency(web, type="both", sparse = FALSE) #Create adjacency matrix from our newly created web
#isSymmetric.matrix(mat) #we should get a full symmetric adjacency matrix since we added the type="both" argument above; this makes sure

V(web)$type <- bipartite.mapping(web)$type #Assign our bipartite mapping to type

V(web)$color <-  NA #Create empty node-propert for color
V(web)$color[V(web)$type==TRUE] <- "red"
V(web)$color[V(web)$type==FALSE] <- "blue"
```


```{r basic plot of igraph web eval=FALSE}
plot(web, layout=layout_as_bipartite(web), edge.width=0.25, vertex.label=NA, vertex.size=1)


plot_data <- dat %>% mutate(
        to = factor(Plant_Species, levels = unique(c(dat$Plant_Species, dat$Frugivore_Species))),
        from = factor(Frugivore_Species, levels = unique(c(dat$Plant_Species, dat$Frugivore_Species))))

plot_data <- dat %>% mutate(
        to = factor(Plant_Species),
        from = factor(Frugivore_Species))

ggplot(plot_data, aes(x = from, y = to)) +
      geom_raster() +
      theme_bw() +
      # Because we need the x and y axis to display every node,
      # not just the nodes that have connections to each other,
      # make sure that ggplot does not drop unused factor levels
      scale_x_discrete(drop = FALSE) +
      scale_y_discrete(drop = FALSE) +
      theme(
        # Rotate the x-axis lables so they are legible
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        # Force the plot into a square aspect ratio
        aspect.ratio = 1,
        # Hide the legend (optional)
        legend.position = "none")
```

# I need a phylogeny for my species

```{r trimming overall BIEN plant trees, eval=FALSE}
plant_list <- unique(dat$Plant_Species)
plant_list <- gsub(" ","_", plant_list)

BIEN_tree <- BIEN_phylogeny_complete(n_phylogenies = 1, seed=1, replicates = 1)
length(BIEN_tree$tip.label)

length(BIEN_matches)
BIEN_matches <- plant_list[plant_list %in% BIEN_tree$tip.label==TRUE]

BIEN_subtree <- ape::keep.tip(BIEN_tree, tip=BIEN_matches)
#pdf(file="BIEN_planttree.pdf")
plot(BIEN_subtree, show.tip.label=FALSE)
```

```{r load Bien subtrees, eval=TRUE}
load("BIEN_subtree.Rda")
```

Now let's get a good frugivore phylogeny
```{r selecting frugivores to query VertLife, eval=FALSE}
Mammal_List <- filter(dat, Frug_Class=="Mammalia") %>% select(., Frugivore_Species) %>% unique()
Bird_List <- filter(dat, Frug_Class=="Aves") %>% select(., Frugivore_Species) %>% unique()

library(clipr)
clipr::write_clip(Mammal_List[,1]) #This way we can copy our species into the VertLife website to request a tree
clipr::write_clip(Bird_List[,1])
```

```{r read and match vertTrees, eval=FALSE}
Mammal_trees <- ape::read.nexus(file="Data.nosync/VertLife_FrugMam/output.nex")

plot(Mammal_trees[1], show.tip.label = FALSE)

Mammal_List <- gsub(" ", "_", Mammal_List[,1])
Mammal_List[Mammal_List %in% Mammal_trees$tree_1642$tip.label==FALSE] #mammals not in our phylogeny
length(Mammal_List)

Bird_trees <- ape::read.nexus(file="Data.nosync/VertLife_FrugBird/output.nex")

plot(Bird_trees[1], show.tip.label = FALSE)

Bird_List <- gsub(" ", "_", Bird_List[,1])
Bird_List[Bird_List %in% Bird_trees$tree_2110$tip.label==FALSE] #mammals not in our phylogeny
length(Bird_List)
length(Bird_List[Bird_List %in% Bird_trees$tree_2110$tip.label==FALSE])
```


```{r load trees, eval=TRUE}
Mammal_trees <- ape::read.nexus(file="Data.nosync/VertLife_FrugMam/output.nex")
Bird_trees <- ape::read.nexus(file="Data.nosync/VertLife_FrugBird/output.nex")
```

```{plot our trees, eval=FALSE}
#pdf(file="Phylogenies.pdf", height=11, width=8.5)

par(mfrow = c(3, 2))
ape::plot.phylo(BIEN_subtree, show.tip.label = FALSE, type="fan", show.node.label = FALSE, main="Plants")
plot(BIEN_subtree, show.tip.label=FALSE)

ape::plot.phylo(Mammal_trees$tree_1642, show.tip.label = FALSE, type="fan", show.node.label = FALSE, main="Mammals")
plot(Mammal_trees$tree_1642, show.tip.label=FALSE)

ape::plot.phylo(Bird_trees$tree_2110, show.tip.label = FALSE, type="fan", show.node.label = FALSE, main="Birds")
plot(Bird_trees$tree_2110, show.tip.label=FALSE)
```


In order to get it working with plug and play, we need to turn our categorical variables into binary ones
```{r factors to binary}
dat <- dplyr::filter_at(dat, vars(Plant_Form, Frugivory_score, Lipid_Score, Fruit_color, Plant_origin), all_vars(!is.na(.))) #make sure we have scores for all our binary variables
dat %<>% dplyr::filter(., Fruit_color !="") #This Remove blank fruit color entries
dat$Plant_origin[dat$Plant_origin !="native"] <- "nonnative" #condense all nonnative plants into one (can revisit later)

#dat$real <- 1 # note that these are real interactions; important when we expand later

dat_old <- dat
dat$Frugivory_score <- paste("FrugScore", dat$Frugivory_score, sep="") #Make value names unambiguous as column names
dat$Lipid_Score <- paste("LipScore", dat$Lipid_Score, sep="")


categories <- list(Forms=unique(dat$Plant_Form), #take the unique entries of our categorical data and save them in a list; (less typing to assign column names based on them below)
                   Frugivory_score=unique(dat$Frugivory_score),
                   Lipid_Score=unique(dat$Lipid_Score),
                   Fruit_color=unique(dat$Fruit_color),
                   Plant_origin=unique(dat$Plant_origin)
                   )
dat_old <- dat

dat <- dat %>% mutate(bin=1) %>% pivot_wider(., names_from = Plant_Form, values_from = bin)
dat <- dat %>% mutate(bin=1) %>% pivot_wider(., names_from = Frugivory_score, values_from = bin)
dat <- dat %>% mutate(bin=1) %>% pivot_wider(., names_from = Lipid_Score, values_from = bin)
dat <- dat %>% mutate(bin=1) %>% pivot_wider(., names_from = Fruit_color, values_from = bin)
dat <- dat %>% mutate(bin=1) %>% pivot_wider(., names_from = Plant_origin, values_from = bin)


dat[, which(colnames(dat)=="liana"):ncol(dat)][is.na(dat[,which(colnames(dat)=="liana"):ncol(dat)])==TRUE] <- 0 #Replae all NA's in our newly created columns with 0

dat <- left_join(dat, dat_old) #add back in our factor columns just in case we want them later
```



#Doing Phylogenetic Decompositions and Adding them to dat

Since we have one tree for the plant species, I just do a singular decomposition and assign the values to the big dataframe. 
```{r add plant phylogeny to dat}
plant_litter <- PVR::PVRdecomp(phy=BIEN_subtree, type="newick", scale=TRUE)

sum((round((plant_litter@Eigen$values)/sum((plant_litter@Eigen$values)),3)*100)[1:4]) 

plant_PhyEig <-data.frame(BIEN_subtree$tip.label, plant_litter@Eigen$vectors[,1:5])
colnames(plant_PhyEig) <- c("Plant_Species", paste(colnames(plant_PhyEig)[2:ncol(plant_PhyEig)], "PlDecomp", sep=""))

plant_PhyEig$Plant_Species <- gsub(pattern="_", replace=" ", x=plant_PhyEig$Plant_Species)

dat <- left_join(dat, plant_PhyEig, by="Plant_Species")
```

Frugivore trees-I'm doing this dumb
```{r eval=FALSE}
frug_tree <- ape::compute.brlen(frug_tree, 1) #set branch lengths to 1
frug_tree <- phytools::force.ultrametric(frug_tree) #force back to being ultrametric (may be bad)
frug_litter <- PVR::PVRdecomp(phy=frug_tree, type="newick", scale=TRUE)

round((frug_litter@Eigen$values)/sum((frug_litter@Eigen$values)),3)*100 #first three vectors contain about 33.8% of variation

frug_PhyEig <-data.frame(frug_tree$tip.label, frug_litter@Eigen$vectors[,1:3])

frug_PhyEig %<>% rename(., "Frugivore_Species"=frug_tree.tip.label)
frug_PhyEig$Frugivore_Species <- gsub("_ott.*","", x=frug_PhyEig$Frugivore_Species)
frug_PhyEig$Frugivore_Species <- gsub(pattern="_", replace=" ", x=frug_PhyEig$Frugivore_Species)
colnames(frug_PhyEig) <- c("Frugivore_Species", "fc1", "fc2", "fc3")

dat <- left_join(dat, frug_PhyEig, by="Frugivore_Species")
```


Now that we're set up for success, let's create our subsets
```{r}
bats <- dplyr::filter(dat, Frug_Group=="Bats")
birds <- dplyr::filter(dat, Frug_Class=="Aves")
```
# Apply Plug and Play to Bats

  
## Phylogeny - Bats
  
```{r Setup bat phylogeny}
frug_litter <- PVR::PVRdecomp(phy=Mammal_trees[[1]], type="newick", scale=TRUE)

#sum((round((frug_litter@Eigen$values)/sum((frug_litter@Eigen$values)),3)*100)[1:3]) #first three vectors contain 66.4% of variation!!!
frug_PhyEig <-data.frame(Mammal_trees[[1]]$tip.label, frug_litter@Eigen$vectors[,1:4]) #take first three vectors and make them a dataframe
colnames(frug_PhyEig) <- c("Frugivore_Species", "fc1", "fc2", "fc3", "fc4")

frug_PhyEig$Frugivore_Species <- gsub(pattern="_", replace=" ", x=frug_PhyEig$Frugivore_Species) 

bats_temp <- left_join(bats, frug_PhyEig, by="Frugivore_Species") #Make a new object that includes our phylo information
bats <- bats_temp
```


  
  

Adding in the Bird Phylogeny  
```{r Setup Bird phylogeny}
frug_litter <- PVR::PVRdecomp(phy=Bird_trees[[5]], type="newick", scale=TRUE)

sum((round((frug_litter@Eigen$values)/sum((frug_litter@Eigen$values)),3)*100)[1:3]) #first three vectors contain 47% of variation
frug_PhyEig <-data.frame(Bird_trees[[5]]$tip.label, frug_litter@Eigen$vectors[,1:3]) #take first three vectors and make them a dataframe
colnames(frug_PhyEig) <- c("Frugivore_Species", "fc1", "fc2", "fc3")

frug_PhyEig$Frugivore_Species <- gsub(pattern="_", replace=" ", x=frug_PhyEig$Frugivore_Species) 

birds_temp <- left_join(birds, frug_PhyEig, by="Frugivore_Species") #Make a new object that includes our phylo information
birds <- birds_temp
```

Latent Link Prediction Methods

Tad's Method
  locSim <- rowMeans(as.matrix(vegan::vegdist(intMat)))	
  pathSim <- rowMeans(as.matrix(vegan::vegdist(t(intMat))))
  simMat <- locSim %*% t(pathSim)	
```{r}
vegstuff <- function(dat){
mat_assym <- as.data.frame.matrix(table(dat$Frugivore_Species, dat$Plant_Species)) #Rows = Frugivores, Columns = Plants
frugSim <- rowMeans(as.matrix(vegan::vegdist(mat_assym)))	 #What's my average distance from all other frugivores (higher means I'm more unique, lower means I'm more similiar)
plantSim <- rowMeans(as.matrix(vegan::vegdist(t(mat_assym)))) #Same thing for plants
simMat <- frugSim %*% t(plantSim)	#uniqueness of the plant * uniqueness of the frugivore
rownames(simMat) <- rownames(mat_assym)
data.frame(frugSim) %>% rownames_to_column("Frugivore_Species")

dat <- data.frame(frugSim) %>% rownames_to_column("Frugivore_Species") %>% left_join(dat, by="Frugivore_Species")
dat_new <- data.frame(plantSim) %>% rownames_to_column("Plant_Species") %>% left_join(dat, by="Plant_Species")
return(dat_new)
}

dat <- vegstuff(dat)
```


```{r}
#bats <- dplyr::filter(dat, Frug_Group=="Bats") %>% dplyr::select(., -frugSim, -plantSim)
bats <- vegstuff(bats)
```

```{r}
#birds <- dplyr::filter(dat, Frug_Class=="Aves") %>% dplyr::select(., -frugSim, -plantSim)
birds <- vegstuff(birds)
```


I've also set up some of the spectral decomposition stuff in another .rmd in the repository.

Now, let's try some random forest models. 

```{r}
woodedWalk <- function(dat, FrugTraits, PlantTraits, class_balancing=FALSE, balance_ratio=3){
  #Set up our data into a fully expanded edgelist
  require(tidyr)
  dat$real <- 1 #make a new column denoting all of these edges are real; important when we expand out to all pairwise connections later
  dat <- dplyr::filter_at(dat, vars(c(FrugTraits, PlantTraits)), all_vars(!is.na(.))) #make sure we have data for all our predictors
  dat <- dplyr::select_at(dat, vars(c(Frugivore_Species, Plant_Species, real, FrugTraits, PlantTraits))) #Select our relevant predictors
  dat %<>% unique(.) #Make sure we only have unique entries
  full_L <- tidyr::expand(dat, Frugivore_Species, Plant_Species) #Expand to include all possible pairwise plant-frugivore interactions.
  full_real <- dplyr::select(dat, Frugivore_Species, Plant_Species, real) %>% left_join(full_L, ., by = c("Frugivore_Species", "Plant_Species")) #notating which of our edges in this expanded data are real 
  full_real$real[is.na(full_real$real)==TRUE] <- 0

  full_real_frugs <- dat[,c("Frugivore_Species", FrugTraits)] %>% unique() %>% left_join(full_real, ., by="Frugivore_Species") #Add in our frugivore traits

  full_real_both <- dat[,c("Plant_Species", PlantTraits)] %>% unique() %>% left_join(full_real_frugs, ., by="Plant_Species") #add in our plant traits
  
  full_real_both$real <- as.factor(full_real_both$real)
  
  if(class_balancing==TRUE){
    # Subsample our edgelist
    ones <- full_real_both %>% dplyr::filter(., real==1)
    #one_IDs <- sample(round(nrow(ones)*0.8), replace=F)
    #testOnes <- ones[-one_IDs,]
    #trainOnes <- ones[one_IDs,]
    
    zeroes <- full_real_both %>% dplyr::filter(., real==0)
    zero_IDs <- sample(x=1:nrow(zeroes), size=nrow(ones)*balance_ratio, replace=F) #Ratio of train to test 1:3
    keptZeroes <- zeroes[zero_IDs,]
    
    #nrow(rbind(trainZeroes, testZeroes, trainOnes, testOnes)) == nrow(full_real_both) #Making sure we're not losing anything
    balanced <- rbind(ones, keptZeroes)
    
    balanced_IDs <- sample(x=1:nrow(balanced),size=round(nrow(balanced)*0.8), replace=F) #subsample out 80% of our edges. 
    test <- balanced[-c(balanced_IDs),] #assign 0% train set
    train <- balanced[balanced_IDs,] #assign 80% train set
  }
    
  if(class_balancing==FALSE){
       # Just Subsample our edgelist as is-don't do any balancing
    sample_IDs <- sample(x=1:nrow(full_real_both), size=round(nrow(full_real_both)*0.8), replace=F) #subsample out 80% of our edges. 
    test <- full_real_both[-sample_IDs,] #assign 80% train set
    train <- full_real_both[sample_IDs,]
  }

  rf<-randomForest::randomForest(real ~. -Frugivore_Species - Plant_Species, data=train, ntree=100)
  trainROC <- roc(train$real,rf$votes[,2])
  predictions <- stats::predict(rf, newdata=test, type="prob")
  test$S <- predictions[,2]
  
  model=list(rfModel=rf,test=test, train=train)
  return(model)
}
```

## Apply Random forest to Bats

Traits - Bats
```{r PnP on Bat Traits}
bats_rf_traits <- woodedWalk(bats, FrugTraits = c("Frug_Body_Mass","Frug_Mean_Gape_Size", "FrugScore1", "FrugScore2", "FrugScore3"), PlantTraits = c("fruit_diameter", "fruit_length", "tree", "palm", "scrub", "yellow", "red", "black", "brown", "green", "LipScore1", "LipScore2"))
#Predictors include continuous variables plus plant form and lipid score


AUC_RF_BatsTraits <- roc(data=bats_rf_traits$test, response=real, predictor=S)
plot(AUC_RF_BatsTraits)
AUC_RF_BatsTraits
```
Bats - Phylogeny
```{r}
bats_rf_plylo <- woodedWalk(bats_temp, FrugTraits = c("fc1", "fc2", "fc3", "fc4"), PlantTraits = c("c1PlDecomp", "c2PlDecomp", "c3PlDecomp", "c4PlDecomp"))

AUC_RF_BatsPhylo <- roc(data=bats_rf_plylo$test, response=real, predictor=S)
plot(AUC_RF_BatsPhylo)
AUC_RF_BatsPhylo
```



### Apply Random Forest to Birds

## Traits - Birds

```{r PnP on Bird Traits}
birds_rf_traits <- woodedWalk(birds, FrugTraits = c("Frug_Body_Mass","Frug_Mean_Gape_Size", "FrugScore1", "FrugScore2", "FrugScore3"), PlantTraits = c("fruit_diameter", "fruit_length", "tree","liana", "palm", "scrub", "yellow", "red", "black", "brown", "green", "LipScore1", "LipScore2", "LipScore3"))#Predictors include continuous variables plus plant form and lipid score


AUC_RF_BirdsTraits <- roc(data=birds_rf_traits$test, response=real, predictor=S)
plot(AUC_RF_BirdsTraits)
AUC_RF_BirdsTraits

```
## Phylogeny - Birds
  
Run it
```{r PnP on Bird Phylogeny}
birds_rf_plylo <- woodedWalk(birds_temp, FrugTraits = c("fc1", "fc2", "fc3"), PlantTraits = c("c1PlDecomp", "c2PlDecomp", "c3PlDecomp", "c4PlDecomp"))
#Predictors include continuous variables plus plant form and lipid score

AUC_RF_BirdsPhylo <- roc(data=birds_rf_plylo$test, response=real, predictor=S)
plot(AUC_RF_BirdsPhylo)
AUC_RF_BirdsPhylo
```
## Latent: Random Forest

```{r}
birds_rf_vegdist <- woodedWalk(birds, FrugTraits = c("frugSim"), PlantTraits = c("plantSim"))

AUC_RF_BirdsVegdist <- roc(data=birds_rf_vegdist$test, response=real, predictor=S)
plot(AUC_RF_BirdsVegdist)
AUC_RF_BirdsVegdist

bats_rf_vegdist <- woodedWalk(bats, FrugTraits = c("frugSim"), PlantTraits = c("plantSim"))

AUC_RF_BatsVegdist <- roc(data=bats_rf_vegdist$test, response=real, predictor=S)
plot(AUC_RF_BatsVegdist)
AUC_RF_BatsVegdist
```




#pdf(file = "Bats_PPEvsRF.pdf", height = 11, width=8.5)
par(mfrow=c(3,2))
plot(AUC_BatsTraits, main="Plug & Play Traits ROC; Bats")
mtext(paste("AUC", round(AUC_BatsTraits$auc, 3), sep="="), side=1)
plot(AUC_RF_BatsTraits, main="Random Forest Traits ROC; Bats")
mtext(paste("AUC", round(AUC_RF_BatsTraits$auc, 3), sep="="), side=1)
plot(AUC_BatsPhylo, main="Plug & Play Phylo ROC; Bats")
mtext(paste("AUC", round(AUC_BatsPhylo$auc, 3), sep="="), side=1)
plot(AUC_RF_BatsPhylo, main="Random Forest Phylo ROC; Bats")
mtext(paste("AUC", round(AUC_RF_BatsPhylo$auc, 3), sep="="), side=1)
plot(AUC_BatsVegdist, main="Plug & Play Vegdist(Latent) ROC; Bats")
mtext(paste("AUC", round(AUC_BatsVegdist$auc, 3), sep="="), side=1)
plot(AUC_RF_BatsVegdist, main="Random Forest Vegdist(Latent) ROC; Bats")
mtext(paste("AUC", round(AUC_RF_BatsVegdist$auc, 3), sep="="), side=1)


#Comparisons
First, put our info types in boxes for ease of use
```{r}
p_phylo <- c("c1PlDecomp", "c2PlDecomp", "c3PlDecomp", "c4PlDecomp")
p_traits <- c("fruit_diameter", "fruit_length", "tree","liana", "palm", "scrub", "yellow", "red", "black", "brown", "green", "LipScore1", "LipScore2", "LipScore3")
p_latent <- c("plantSim")

f_phylo <- c("fc1", "fc2", "fc3")
f_traits <- c("Frug_Body_Mass","Frug_Mean_Gape_Size", "FrugScore1", "FrugScore2", "FrugScore3")
f_latent <- c("frugSim")
```


Starting with random forest.

RF Phylo + Trait.
```{r}
birds_rf_PhyTraits <- woodedWalk(birds_temp, FrugTraits = c(f_phylo, f_traits), PlantTraits = c(p_phylo, p_traits))

birds_rf_TraitsLatent <- woodedWalk(birds, FrugTraits = c(f_latent, f_traits), PlantTraits = c(p_latent, p_traits))

birds_rf_PhyLatent <- woodedWalk(birds, FrugTraits = c(f_latent, f_traits), PlantTraits = c(p_latent, p_traits))

birds_rf_Trio <- woodedWalk(birds, FrugTraits = c(f_phylo, f_latent, f_traits), PlantTraits = c(p_phylo, p_latent, p_traits))

AUC_birds_rf_TraitsLatent <- roc(data=birds_rf_TraitsLatent$test, response=real, predictor=S)
plot(AUC_birds_rf_TraitsLatent)
AUC_birds_rf_TraitsLatent

AUC_birds_rf_PhyTraits <- roc(data=birds_rf_PhyTraits$test, response=real, predictor=S)
plot(AUC_birds_rf_PhyTraits)
AUC_birds_rf_PhyTraits

AUC_birds_rf_PhyLatent <- roc(data=birds_rf_PhyLatent$test, response=real, predictor=S)
plot(AUC_birds_rf_PhyLatent)
AUC_birds_rf_PhyLatent

AUC_birds_rf_Trio <- roc(data=birds_rf_Trio$test, response=real, predictor=S)
plot(AUC_birds_rf_Trio)
AUC_birds_rf_Trio
```

```{r}
#pdf(file = "Birds_RF_Combos.pdf", height = 11, width=8.5)
par(mfrow=c(3,2))
plot(AUC_birds_rf_TraitsLatent, main="RF Traits & Latent ROC; Birds")
mtext(paste("AUC", round(AUC_birds_rf_TraitsLatent$auc, 3), sep="="), side=1)

plot(AUC_birds_rf_PhyTraits, main="RF Phylo & Traits ROC; Birds")
mtext(paste("AUC", round(AUC_birds_rf_PhyTraits$auc, 3), sep="="), side=1)

plot(AUC_birds_rf_PhyLatent, main="RF Phylo & Latent ROC; Birds")
mtext(paste("AUC", round(AUC_birds_rf_PhyLatent$auc, 3), sep="="), side=1)

plot(AUC_birds_rf_Trio, main="RF Trio ROC; Birds")
mtext(paste("AUC", round(AUC_birds_rf_Trio$auc, 3), sep="="), side=1)
```


```{r}
bats_rf_PhyTraits <- woodedWalk(bats, FrugTraits = c(f_phylo, f_traits), PlantTraits = c(p_phylo, p_traits))

bats_rf_TraitsLatent <- woodedWalk(bats, FrugTraits = c(f_latent, f_traits), PlantTraits = c(p_latent, p_traits))

bats_rf_PhyLatent <- woodedWalk(bats, FrugTraits = c(f_latent, f_traits), PlantTraits = c(p_latent, p_traits))

bats_rf_Trio <- woodedWalk(bats, FrugTraits = c(f_phylo, f_latent, f_traits), PlantTraits = c(p_phylo, p_latent, p_traits))

AUC_bats_rf_TraitsLatent <- roc(data=bats_rf_TraitsLatent$test, response=real, predictor=S)
plot(AUC_bats_rf_TraitsLatent)
AUC_bats_rf_TraitsLatent

AUC_bats_rf_PhyTraits <- roc(data=bats_rf_PhyTraits$test, response=real, predictor=S)
plot(AUC_bats_rf_PhyTraits)
AUC_bats_rf_PhyTraits

AUC_bats_rf_PhyLatent <- roc(data=bats_rf_PhyLatent$test, response=real, predictor=S)
plot(AUC_bats_rf_PhyLatent)
AUC_bats_rf_PhyLatent

AUC_bats_rf_Trio <- roc(data=bats_rf_Trio$test, response=real, predictor=S)
AUC_bats_rf_Trio
AUC_bats_rf_Trio
```

```{r}
#pdf(file = "Bats_RF_Combos.pdf", height = 11, width=8.5)
par(mfrow=c(3,2))
plot(AUC_bats_rf_TraitsLatent, main="RF Traits & Latent ROC; bats")
mtext(paste("AUC", round(AUC_bats_rf_TraitsLatent$auc, 3), sep="="), side=1)

plot(AUC_bats_rf_PhyTraits, main="RF Phylo & Traits ROC; bats")
mtext(paste("AUC", round(AUC_bats_rf_PhyTraits$auc, 3), sep="="), side=1)

plot(AUC_bats_rf_PhyLatent, main="RF Phylo & Latent ROC; bats")
mtext(paste("AUC", round(AUC_bats_rf_PhyLatent$auc, 3), sep="="), side=1)

plot(AUC_bats_rf_Trio, main="RF Trio ROC; bats")
mtext(paste("AUC", round(AUC_bats_rf_Trio$auc, 3), sep="="), side=1)
```


```{r}
getAccurate <- function(rfob, threshold=0.5){
  require(ROCR); require(hmeasure)
  
  auc <- roc(data=rfob$test, response=real, predictor=S)$auc
  hmeas <- hmeasure::HMeasure(true.class=rfob$test$real, scores=rfob$test$S)$metrics$H
	mrse <- sqrt(sum((rfob$test$S-as.numeric(rfob$test$real))^2))
	acc <- max(unlist(ROCR::performance(ROCR::prediction(rfob$test$S,rfob$test$real),'acc')@y.values))
	
  return(list(
		acc=acc,
		auc=auc,
		mrse=mrse,
		H=hmeas,
		m=deparse(substitute(rfob)) #return the name of your rfob input
  ))
}

birds_rf_accuracy <- data.frame(rbind(unlist(getAccurate(birds_rf_traits)),
      unlist(getAccurate(birds_rf_vegdist)),
      unlist(getAccurate(birds_rf_plylo)),
      unlist(getAccurate(birds_rf_PhyLatent)),
      unlist(getAccurate(birds_rf_PhyTraits)),
      unlist(getAccurate(birds_rf_TraitsLatent)),
      unlist(getAccurate(birds_rf_Trio))))



makeComPlot <- function(temp){
temp$acc %<>% as.numeric(.)
temp$auc %<>% as.numeric(.)
temp$H %<>% as.numeric(.)
temp$mrse %<>% as.numeric(.)

temp$acc <- temp$acc/max(temp$acc)
temp$auc <- temp$auc/max(temp$auc)
temp$mrse <- temp$mrse/max(temp$mrse)
temp$H <- temp$H/max(temp$H)


pivot_longer(temp, 1:4, names_to = "stat") %>% ggplot(data=., aes(fill=stat, y=value, x=m)) + 
    geom_bar(position="dodge", stat="identity")+xlab("Model")+ylab("Relative performance")+ggtitle("Bird Random Forest Link Prediction Performance Metrics (Value/Maximum)")+theme(axis.text = element_text(size=12))
}

```


```{r}
ggplot(data=birds_rf_accuracy, aes(x=m, y=H))+geom_bar(aes(y=H), stat="identity", position=position_dodge())+geom_bar(aes(y=acc), stat="identity", position=position_dodge(0.9))

birds_rf_accuracy[,1:4] <- lapply(birds_rf_accuracy[,1:4], as.numeric)
 
# Grouped

birds_rf_accuracy$acc <- birds_rf_accuracy$acc/max(birds_rf_accuracy$acc)
birds_rf_accuracy$auc <- birds_rf_accuracy$auc/max(birds_rf_accuracy$auc)
birds_rf_accuracy$mrse <- birds_rf_accuracy$mrse/max(birds_rf_accuracy$mrse)
birds_rf_accuracy$H <- birds_rf_accuracy$H/max(birds_rf_accuracy$H)
pivot_longer(birds_rf_accuracy, 1:4, names_to = "stat")



#pdf(file="bird_Rf_relcomp.pdf", height=8.5, width=11)
pivot_longer(birds_rf_accuracy, 1:4, names_to = "stat") %>% ggplot(data=., aes(fill=stat, y=value, x=m)) + 
    geom_bar(position="dodge", stat="identity")+xlab("Model")+ylab("Relative performance")+ggtitle("Bird Random Forest Link Prediction Performance Metrics (Value/Maximum)")+theme(axis.text = element_text(size=12))



bats_rf_accuracy <- data.frame(rbind(unlist(getAccurate(bats_rf_traits)),
      unlist(getAccurate(bats_rf_vegdist)),
      unlist(getAccurate(bats_rf_plylo)),
      unlist(getAccurate(bats_rf_PhyLatent)),
      unlist(getAccurate(bats_rf_PhyTraits)),
      unlist(getAccurate(bats_rf_TraitsLatent)),
      unlist(getAccurate(bats_rf_Trio))))


#pdf(file="bat_Rf_relcomp.pdf", height=8.5, width=11)
makeComPlot(bats_rf_accuracy)
```

Now, I'm redoing this after reducing the class imbalances in the training data

```{r}
birds_rebalancedRF_Phy <- woodedWalk(birds, FrugTraits = c(f_phylo), PlantTraits = c(p_phylo), class_balancing = TRUE)

birds_rebalancedRF_Traits <- woodedWalk(birds, FrugTraits = c(f_traits), PlantTraits = c(p_traits), class_balancing = TRUE)

birds_rebalancedRF_Latent <- woodedWalk(birds, FrugTraits = c(f_latent), PlantTraits = c(p_latent), class_balancing = TRUE)

birds_rebalancedRF_PhyTraits <- woodedWalk(birds, FrugTraits = c(f_phylo, f_traits), PlantTraits = c(p_phylo, p_traits), class_balancing = TRUE)

birds_rebalancedRF_TraitsLatent <- woodedWalk(birds, FrugTraits = c(f_latent, f_traits), PlantTraits = c(p_latent, p_traits), class_balancing = TRUE)

birds_rebalancedRF_PhyLatent <- woodedWalk(birds, FrugTraits = c(f_latent, f_phylo), PlantTraits = c(p_latent, p_phylo), class_balancing = TRUE)

birds_rebalancedRF_Trio <- woodedWalk(birds, FrugTraits = c(f_phylo, f_latent, f_traits), PlantTraits = c(p_phylo, p_latent, p_traits), class_balancing = TRUE)

AUC_birds_rebalancedRF_Phy <- roc(data=birds_rebalancedRF_Phy$test, response=real, predictor=S)
plot(AUC_birds_rebalancedRF_Phy)
AUC_birds_rebalancedRF_Phy


AUC_birds_rebalancedRF_Traits <- roc(data=birds_rebalancedRF_Traits$test, response=real, predictor=S)
plot(AUC_birds_rebalancedRF_Traits)
AUC_birds_rebalancedRF_Traits

AUC_birds_rebalancedRF_Latent <- roc(data=birds_rebalancedRF_Latent$test, response=real, predictor=S)
plot(AUC_birds_rebalancedRF_Latent)
AUC_birds_rebalancedRF_Latent

AUC_birds_rebalancedRF_PhyTraits <- roc(data=birds_rebalancedRF_PhyTraits$test, response=real, predictor=S)
plot(AUC_birds_rebalancedRF_PhyTraits)
AUC_birds_rebalancedRF_PhyTraits

AUC_birds_rebalancedRF_TraitsLatent <- roc(data=birds_rebalancedRF_TraitsLatent$test, response=real, predictor=S)
plot(AUC_birds_rebalancedRF_TraitsLatent)
AUC_birds_rebalancedRF_TraitsLatent

AUC_birds_rebalancedRF_PhyLatent <- roc(data=birds_rebalancedRF_PhyLatent$test, response=real, predictor=S)
plot(AUC_birds_rebalancedRF_PhyLatent)
AUC_birds_rebalancedRF_PhyLatent

AUC_birds_rebalancedRF_Trio <- roc(data=birds_rebalancedRF_Trio$test, response=real, predictor=S)
plot(AUC_birds_rebalancedRF_Trio)
AUC_birds_rebalancedRF_Trio



birds_rebalancedRF_accuracy <- data.frame(rbind(unlist(getAccurate(birds_rebalancedRF_Phy)),
      unlist(getAccurate(birds_rebalancedRF_Traits)),
      unlist(getAccurate(birds_rebalancedRF_Latent)),
      unlist(getAccurate(birds_rebalancedRF_PhyTraits)),
      unlist(getAccurate(birds_rebalancedRF_TraitsLatent)),
      unlist(getAccurate(birds_rebalancedRF_PhyLatent)),
      unlist(getAccurate(birds_rebalancedRF_Trio))))

birds_rf_accuracy

birds_rebalancedRF_accuracy

```



```{r}
bats_rebalancedRF_Phy <- woodedWalk(bats, FrugTraits = c(f_phylo), PlantTraits = c(p_phylo), class_balancing = TRUE)

bats_rebalancedRF_Traits <- woodedWalk(bats, FrugTraits = c(f_traits), PlantTraits = c(p_traits), class_balancing = TRUE)

bats_rebalancedRF_Latent <- woodedWalk(bats, FrugTraits = c(f_latent), PlantTraits = c(p_latent), class_balancing = TRUE)

bats_rebalancedRF_PhyTraits <- woodedWalk(bats, FrugTraits = c(f_phylo, f_traits), PlantTraits = c(p_phylo, p_traits), class_balancing = TRUE)

bats_rebalancedRF_TraitsLatent <- woodedWalk(bats, FrugTraits = c(f_latent, f_traits), PlantTraits = c(p_latent, p_traits), class_balancing = TRUE)

bats_rebalancedRF_PhyLatent <- woodedWalk(bats, FrugTraits = c(f_latent, f_phylo), PlantTraits = c(p_latent, p_phylo), class_balancing = TRUE)

bats_rebalancedRF_Trio <- woodedWalk(bats, FrugTraits = c(f_phylo, f_latent, f_traits), PlantTraits = c(p_phylo, p_latent, p_traits), class_balancing = TRUE)

AUC_bats_rebalancedRF_Phy <- roc(data=bats_rebalancedRF_Phy$test, response=real, predictor=S)
plot(AUC_bats_rebalancedRF_Phy)
AUC_bats_rebalancedRF_Phy

AUC_bats_rebalancedRF_Traits <- roc(data=bats_rebalancedRF_Traits$test, response=real, predictor=S)
plot(AUC_bats_rebalancedRF_Traits)
AUC_bats_rebalancedRF_Traits

AUC_bats_rebalancedRF_Latent <- roc(data=bats_rebalancedRF_Latent$test, response=real, predictor=S)
plot(AUC_bats_rebalancedRF_Latent)
AUC_bats_rebalancedRF_Latent

AUC_bats_rebalancedRF_PhyTraits <- roc(data=bats_rebalancedRF_PhyTraits$test, response=real, predictor=S)
plot(AUC_bats_rebalancedRF_PhyTraits)
AUC_bats_rebalancedRF_PhyTraits

AUC_bats_rebalancedRF_TraitsLatent <- roc(data=bats_rebalancedRF_TraitsLatent$test, response=real, predictor=S)
plot(AUC_bats_rebalancedRF_TraitsLatent)
AUC_bats_rebalancedRF_TraitsLatent

AUC_bats_rebalancedRF_PhyLatent <- roc(data=bats_rebalancedRF_PhyLatent$test, response=real, predictor=S)
plot(AUC_bats_rebalancedRF_PhyLatent)
AUC_bats_rebalancedRF_PhyLatent

AUC_bats_rebalancedRF_Trio <- roc(data=bats_rebalancedRF_Trio$test, response=real, predictor=S)
plot(AUC_bats_rebalancedRF_Trio)
AUC_bats_rebalancedRF_Trio
```

Adding in SVD measures!


```{r}
svData <- function(dat){
mat_assym <- as.data.frame.matrix(table(dat$Frugivore_Species, dat$Plant_Species)) #Rows = Frugivores, Columns = Plants

decomp <- svd(mat_assym)#run svd decomposition
u <- decomp$u #extract first 3 axis
v <- data.frame(decomp$v) #I think this is the right dimension?

plantsSVD <-data.frame(colnames(mat_assym), v[,1], v[,2], v[,3])
colnames(plantsSVD) <- c("Plant_Species", "Psvd1", "Psvd2", "Psvd3")
frugSVD <- data.frame(colnames(t(mat_assym)), u[,1], u[,2], u[,3])
colnames(frugSVD) <- c("Frugivore_Species", "Fsvd1", "Fsvd2", "Fsvd3")

dat_wP <- left_join(dat, plantsSVD, by="Plant_Species")
dat_new <- left_join(dat_wP, frugSVD, by="Frugivore_Species")

return(dat_new)
}

birds <- svData(birds)
bats <- svData(bats)
```


```{r}
#p_phylo <- c("c1PlDecomp", "c2PlDecomp", "c3PlDecomp", "c4PlDecomp")
#p_traits <- c("fruit_diameter", "fruit_length", "tree","liana", "palm", "scrub", "yellow", "red", "black", "brown", "green", "LipScore1", "LipScore2", "LipScore3")
#p_latent <- c("plantSim", "Psvd1", "Psvd2", "Psvd3") #Both methods together
p_latent <- c("Psvd1", "Psvd2", "Psvd3")

#f_phylo <- c("fc1", "fc2", "fc3")
#f_traits <- c("Frug_Body_Mass","Frug_Mean_Gape_Size", "FrugScore1", "FrugScore2", "FrugScore3")
#f_latent <- c("frugSim", "Fsvd1", "Fsvd2", "Fsvd3")
f_latent <- c("Fsvd1", "Fsvd2", "Fsvd3")


birds_rebalancedRF_Latent <- woodedWalk(birds, FrugTraits = c(f_latent), PlantTraits = c(p_latent), class_balancing = TRUE)

birds_rebalancedRF_TraitsLatent <- woodedWalk(birds, FrugTraits = c(f_latent, f_traits), PlantTraits = c(p_latent, p_traits), class_balancing = TRUE)

birds_rebalancedRF_PhyLatent <- woodedWalk(birds, FrugTraits = c(f_latent, f_phylo), PlantTraits = c(p_latent, p_phylo), class_balancing = TRUE)

birds_rebalancedRF_Trio <- woodedWalk(birds, FrugTraits = c(f_phylo, f_latent, f_traits), PlantTraits = c(p_phylo, p_latent, p_traits), class_balancing = TRUE)

AUC_birds_rebalancedRF_Latent_SVD <- roc(data=birds_rebalancedRF_Latent$test, response=real, predictor=S)
plot(AUC_birds_rebalancedRF_Latent_SVD)
AUC_birds_rebalancedRF_Latent_SVD


AUC_birds_rebalancedRF_TraitsLatent_SVD <- roc(data=birds_rebalancedRF_TraitsLatent$test, response=real, predictor=S)
plot(AUC_birds_rebalancedRF_TraitsLatent_SVD)
AUC_birds_rebalancedRF_TraitsLatent_SVD

AUC_birds_rebalancedRF_PhyLatent_SVD <- roc(data=birds_rebalancedRF_PhyLatent$test, response=real, predictor=S)
plot(AUC_birds_rebalancedRF_PhyLatent_SVD)
AUC_birds_rebalancedRF_PhyLatent_SVD

AUC_birds_rebalancedRF_Trio_SVD <- roc(data=birds_rebalancedRF_Trio$test, response=real, predictor=S)
plot(AUC_birds_rebalancedRF_Trio_SVD)
AUC_birds_rebalancedRF_Trio_SVD
```
Making some more plots



```{r}
birds_rebalancedRF_accuracy <- data.frame(rbind(unlist(getAccurate(birds_rebalancedRF_Phy)),
      unlist(getAccurate(birds_rebalancedRF_Traits)),
      unlist(getAccurate(birds_rebalancedRF_Latent)),
      unlist(getAccurate(birds_rebalancedRF_PhyTraits)),
      unlist(getAccurate(birds_rebalancedRF_TraitsLatent)),
      unlist(getAccurate(birds_rebalancedRF_PhyLatent)),
      unlist(getAccurate(birds_rebalancedRF_Trio))))
```

