---
title: "Frugivory"
author: "Grant Foster"
date: "9/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(magrittr)
library(igraph)
```


```{r}
dat <- read.csv("ATLANTIC_frugivory.csv")
```

```{r}
uniq<- dat %>% select(., Frugivore_Species, Plant_Species) %>% unique()

web <- data.frame(
    from = uniq$Frugivore_Species,
    to = uniq$Plant_Species
  ) %>% graph_from_data_frame(directed=FALSE)

mat <-get.adjacency(web, type="both", sparse = FALSE)
isSymmetric.matrix(mat)

V(web)$type <- bipartite.mapping(web)$type
igraph::is.bipartite(web)
plot(web, layout=layout_as_bipartite(web))

layout_as_bipartite(web)

```

```{r eval=FALSE}
V(web)$type
V(web)$color <-  NA
V(web)$color[V(web)$type==TRUE] <- "red"
V(web)$color[V(web)$type==FALSE] <- "blue"

plot(web, layout=layout_as_bipartite(web), edge.width=0.25, vertex.label=NA, vertex.size=1)


plot_data <- dat %>% mutate(
        to = factor(Plant_Species, levels = unique(c(dat$Plant_Species, dat$Frugivore_Species))),
        from = factor(Frugivore_Species, levels = unique(c(dat$Plant_Species, dat$Frugivore_Species))))

plot_data <- dat %>% mutate(
        to = factor(Plant_Species),
        from = factor(Frugivore_Species))

ggplot(plot_data, aes(x = from, y = to)) +
      geom_raster() +
      theme_bw() +
      # Because we need the x and y axis to display every node,
      # not just the nodes that have connections to each other,
      # make sure that ggplot does not drop unused factor levels
      scale_x_discrete(drop = FALSE) +
      scale_y_discrete(drop = FALSE) +
      theme(
        # Rotate the x-axis lables so they are legible
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        # Force the plot into a square aspect ratio
        aspect.ratio = 1,
        # Hide the legend (optional)
        legend.position = "none")

```
I need a phylogeny for my species

The frugivore one I get from tree of life is ok, but the plant one kind of sucks-huge polytomies at really pretty high taxonomic levels. 

## Messing with Phylogeneti Trees
```{r eval=FALSE}
library(rotl)
plant_names <- rotl::tnrs_match_names(unique(dat$Plant_Species))
plant_names %<>% filter(., approximate_match ==FALSE) #Only keep correct matches
plant_names %<>% filter(., flags != "incertae_sedis_inherited")
plant_names %<>% filter(., flags != "hybrid, sibling_higher")


dat %<>% filter(., Plant_Species %in% plant_names$unique_name) 


tree <- ape::read.nexus("Jansenss_2020_AngiTree.tre")

plot(tree, type="fan")

dat$PlantName_Under <- gsub(" ","_", dat$Plant_Species)

table(unique(dat$PlantName_Under) %in% tree$tip.label)

plant_tree <- rotl::tol_induced_subtree(ott_ids = plant_names$ott_id)


ape::plot.phylo(plant_tree, show.tip.label = FALSE, type="fan", show.node.label = FALSE)
plot(plant_tree, show.tip.label=FALSE)

frug_names <- rotl::tnrs_match_names(unique(dat$Frugivore_Species))
frug_names %<>% filter(., approximate_match ==FALSE) #Only keep correct matches
frug_names %<>% filter(., flags != "incertae_sedis_inherited")
frug_names %<>% filter(., flags != "hybrid, sibling_higher")

frug_tree <- rotl::tol_induced_subtree(ott_ids = frug_names$ott_id)

ape::plot.phylo(frug_tree, show.tip.label = FALSE, type="fan", show.node.label = FALSE)
plot(frug_tree, show.tip.label=FALSE)
```


Making a smaller example

```{r}
set.seed(24601)

test_web <- igraph::delete.edges(web, sample(1:length(E(web)), 3658, replace = FALSE)) #remove 50% of our data
#plot(web, layout=layout_as_bipartite(web))
plot(test_web, layout=layout_as_bipartite(test_web))


test_adj <- get.adjacency(test_web, type="both", sparse = FALSE)

#lattice::levelplot(test_adj)
isSymmetric.matrix(test_adj)
```

Creating link prediction function

Note: If I make a strictly upper triangular matrix (non-symmetric), it's eigenvalues will all be 0. So we need to see a symmetric matrix. 

We want ideally to minimize the Sigma[f(delta_kk)-U^T * A' * Uk]^2, essentially a sum of squares.
```{r}
U <-eigen(mat)$vectors
L <- eigen(mat)$values
table(mat)
table(mat)/sum(table(mat))


# A = U*L*U^T
test <- U %*% diag(L) %*% t(U)
test <- round(test, digits=3) #Rounding errors make our 0's into super small values
table(test==mat) #This should be true

#A' = F(A) = U*F(L)*U^T, where F is our kernel function relating our training set (A) to test set (A')
#We want to minimize |F(A)-A'|, where

U_train <- eigen(test_adj)$vectors
L_train <- eigen(test_adj)$values

target_V <- t(U_train) %*% mat %*% U_train
target_V <- diag(target_V)

plot(x=L_train, y=target_V, col="dodger blue", pch=16)

```

The first kernel that we use is odd path counting. Here we look at weighted sum of several odd powers of the original adjacency matrix (recall that k-th power of an adjacency matrix refers to numbers of k-hop neighbours from every node). The higher the value, the more likely it is for two vertices to be connected.

I think this is mostly wrong. I hope to fix it in the future.

```{r}
library(expm)

odd_kernal <- function(input, max_path){
  index <- seq(from=1, to=max_path, by=2)
  #index <- index[-1]
  output <- 0
  for(i in index){
    #print(i)
    output <- output + (test_adj %^% i)
  }
  return(output)
}

odd_paths_V1 <- t(U_train) %*% odd_kernal(input=test_adj, max_path = 1) %*% U_train
odd_paths_pred1 <- diag(odd_paths_V1)

odd_paths_V3 <- t(U_train) %*% odd_kernal(input=test_adj, max_path = 3) %*% U_train
odd_paths_pred3 <- diag(odd_paths_V3)

odd_paths_V5 <- t(U_train) %*% odd_kernal(input=test_adj, max_path = 5) %*% U_train
odd_paths_pred5 <- diag(odd_paths_V5)

odd_paths_V7 <- t(U_train) %*% odd_kernal(input=test_adj, max_path = 7) %*% U_train
odd_paths_pred7 <- diag(odd_paths_V7)

odd_paths_V9 <- t(U_train) %*% odd_kernal(input=test_adj, max_path = 9) %*% U_train
odd_paths_pred9 <- diag(odd_paths_V9)

plot(x=L_train, y=target_V, col="dodger blue", pch=16)
points(x=L_train, y=odd_paths_pred1, col="red", pch=16)
points(x=L_train, y=odd_paths_pred3, col="orange", pch=16)
points(x=L_train, y=odd_paths_pred5, col="green", pch=16)
points(x=L_train, y=odd_paths_pred7, col="purple", pch=16)
points(x=L_train, y=odd_paths_pred9, col="navy", pch=16)
legend(1, 1, legend=c("1 (base)", "3", "5", "7", "9"),
       col=c("red", "orange", "green", "purple", "navy"), pch=16, cex=0.8)

all.equal(dim(odd_paths_V1), dim(mat))

edgelist <- data.frame(matrix(mat, dimnames=list(t(outer(colnames(mat), rownames(mat), FUN=paste)), NULL)))
colnames(edgelist) <- "truth"
edgelist <- rownames_to_column(edgelist, var="link")


colnames(odd_paths_V1) <- colnames(mat)
rownames(odd_paths_V1) <- rownames(mat)
odd_paths_1L <- data.frame(matrix(odd_paths_V1, dimnames=list(t(outer(colnames(odd_paths_V1), rownames(odd_paths_V1), FUN=paste)), NULL)))
colnames(odd_paths_1L) <- "Path1"
odd_paths_1L <- rownames_to_column(odd_paths_1L, var="link")

colnames(odd_paths_V3) <- colnames(mat)
rownames(odd_paths_V3) <- rownames(mat)
odd_paths_3L <- data.frame(matrix(odd_paths_V3, dimnames=list(t(outer(colnames(odd_paths_V3), rownames(odd_paths_V3), FUN=paste)), NULL)))
colnames(odd_paths_3L) <- "Path3"
odd_paths_3L <- rownames_to_column(odd_paths_3L, var="link")

ROC_data <- left_join(edgelist, odd_paths_1L, by="link")
ROC_data <- left_join(ROC_data, odd_paths_3L, by="link")

simple_roc <- function(labels, scores){
  labels <- labels[order(scores, decreasing=TRUE)]
  data.frame(TPR=cumsum(labels)/sum(labels), FPR=cumsum(!labels)/sum(!labels), labels)
}

temp <- simple_roc(labels=ROC_data$truth, scores=ROC_data$Path1)
temp2 <- simple_roc(labels=ROC_data$truth, scores=ROC_data$Path3)
plot(x=temp$FPR, y=temp$TPR, type="l", col="blue")
lines(x=temp2$FPR, y=temp2$TPR, type="l", col="red")

library(pROC)


AUC_Odd1 <- roc(response=ROC_data$truth, predictor=ROC_data$Path1)
AUC_Odd3 <- roc(response=ROC_data$truth, predictor=ROC_data$Path3)

auc(AUC_Odd1)
plot(AUC_Odd1)

auc(AUC_Odd3)
plot(AUC_Odd3)

```


Step back: Let's predict based on traits. First, I'll start with a pretty simple trait matching approach.
```{r}
cor.test(x=dat$fruit_diameter, y=dat$Frug_Mean_Gape_Size) 
summary(lm(dat$Frug_Mean_Gape_Size~dat$fruit_diameter))

#Alright, these things (as we'd guess) are highly correlated. Let's use them to predict links.

test_long <- data.frame(get.edgelist(test_web))
colnames(test_long) <- c("Frugivore_Species", "Plant_Species")

test_dat <- left_join(test_long, dat, by=c("Frugivore_Species", "Plant_Species"))


#Perform the regression on our test data
test_dat %<>% filter(., is.na(Frug_Mean_Gape_Size)==FALSE & is.na(fruit_diameter)==FALSE)
Di_Gape_lm <- lm(Frug_Mean_Gape_Size~fruit_diameter, data=test_dat)
summary(Di_Gape_lm)

test_dat$Di_Gape_SSQ <- Di_Gape_lm$residuals^2 #This is only doing it on the test version; not the full dataset!

dat %<>% filter(., is.na(Frug_Mean_Gape_Size)==FALSE & is.na(fruit_diameter)==FALSE)
dat %<>% filter(., is.na(Frug_Mean_Gape_Size)==FALSE & is.na(fruit_diameter)==FALSE)
dat %<>% filter(., is.na(Frugivory_score)==FALSE & is.na(fruit_diameter)==FALSE)
dat %<>% filter(., is.na(fruit_length)==FALSE & is.na(fruit_diameter)==FALSE)

dat$pred_gape <- stats::predict(object=Di_Gape_lm, newdata=data.frame(dat))

dat$Di_Gape_SSQ <- (dat$pred_gape - dat$Frug_Mean_Gape_Size)^2 #This is for all the real interactions; we need all 

full_L <- tidyr::expand(dat, Frugivore_Species, Plant_Species)

full_L <-  dat %>% select(., Frugivore_Species, Frug_Mean_Gape_Size, Frug_Class, Frug_Order, Frug_Family, Frug_Genus, Frug_Epitetus, Frug_Group, Frug_Body_Mass, Frugivory_score, Frug_Migration_status, Frug_IUCN, Frug_Population_Trend) %>% unique() %>% left_join(full_L, ., by="Frugivore_Species")

full_L <-  dat %>% select(., Plant_Species, fruit_diameter, fruit_length, Fruit_color, Plant_distribution, Plant_origin, Plant_Form, Plants_IUCN) %>% unique() %>% left_join(full_L, ., by="Plant_Species")

full_L$pred_gape <- stats::predict(object=Di_Gape_lm, newdata=data.frame(full_L))
full_L$Di_Gape_SSQ <- (full_L$pred_gape - full_L$Frug_Mean_Gape_Size)^2

full_L <- dat %>% select(., Frugivore_Species, Plant_Species) %>% mutate(., real=1) %>% left_join(full_L, ., by=c("Frugivore_Species", "Plant_Species"))

full_L$real[is.na(full_L$real)==TRUE] <- 0

AUC_gape_fruit <- roc(data=data.frame(full_L), response=real, predictor=Di_Gape_SSQ)
AUC_gape_fruit
plot(AUC_gape_fruit)
```

That was a pretty ham-fisted linear approach. How should we go about this otherwise? At first I thought logistic regression might be the best way to go, but realized the general style doens't care about matching traits, just their values. 

What could be a good approach is to do a generalized linear model. 
```{r}

pointsobject <- full_L %>% dplyr::select(., Frugivore_Species, Plant_Species, real, Frug_Mean_Gape_Size, Frug_Body_Mass, Frugivory_score, fruit_diameter, fruit_length, Fruit_color)

p <- pointsobject[(pointsobject[,'real'] == 1), 4:(ncol(pointsobject)-1)]
bgrd <- pointsobject[,4:(ncol(pointsobject)-1)]

#Recreating the plug and play approach

  mean.p <- colMeans(p, na.rm = TRUE)         # estimated mean of presence points
  mean.bgrd <- colMeans(bgrd)   # estimated mean of background points
  sigma.p <- stats::cov(p)             # estimated covariance of presence points
  sigma.bgrd <- stats::cov(bgrd)       # estimated covariance of background points
    
  require(corpcor)
    # robust estimated covariance of presence points
    p.est <- as.numeric(cov.shrink(p, verbose= TRUE))  
    sigma.p <- matrix(as.numeric(p.est), nrow=sqrt(length(p.est)))  #make it into a square matrix?
    # robust estimated covariance of background points
    bgrd.est <- as.numeric(cov.shrink(bgrd, verbose = TRUE))    
    sigma.bgrd <- matrix(as.numeric(bgrd.est), nrow=sqrt(length(bgrd.est))) # reformat

    model=list(mean.p=mean.p, sigma.p=sigma.p, mean.bgrd=mean.bgrd, sigma.bgrd=sigma.bgrd)  

    
predict.pp.gauss <- function(model, x){
  require(mvtnorm)
  f1 <- dmvnorm(x, mean=model$mean.p, sigma=model$sigma.p, log=TRUE)
  f0 <- dmvnorm(x, mean=model$mean.bgrd, sigma=model$sigma.bgrd, log=TRUE)
  return(S=exp(f1-f0))
} 

pp.kde <- function(p, bgrd){
  require(np)
  f1 <- npudens(p, bwmethod='normal-reference')
  f0 <- npudens(bgrd, bwmethod='normal-reference')
  return(model=list(f1=f1,f0=f0))
}

predict.pp.kde <- function(model, x){
  require(np)
  f1 <- fitted(npudens(bws=model$f1$bws, edat=data.frame(x), bwmethod='normal-reference'))
  f0 <- fitted(npudens(bws=model$f0$bws, edat=data.frame(x), bwmethod='normal-reference'))
  return(S=f1/f0)
}

red.preds.gauss <- predict.pp.gauss(model, x=bgrd)
full_L$suit.gauss <- red.preds.gauss

stuff <- pp.kde(p, bgrd)

str(stuff)
red.preds.kde <- predict.pp.kde(stuff, bgrd)


full_L$suit.kde <- red.preds.kde

AUC_PP <- roc(data=data.frame(full_L), response=real, predictor=suit)
AUC_PP
plot(AUC_PP)
```
Simplifying the problem: for both complexity and runtime's sake
Note:frugivory is not actually meaningful, but it's still in. 

14 bat species, 47 plant species.
```{r}
#Make long form of just bat intearactions.
bats <- dplyr::filter(dat, Frug_Group=="Bats")
batsL <- tidyr::expand(bats, Frugivore_Species, Plant_Species)

#Left_Join in our trait data and denote real entries
batsL <-  bats %>% select(., Frugivore_Species, Frug_Mean_Gape_Size, Frug_Body_Mass) %>% unique() %>% left_join(batsL, ., by="Frugivore_Species")
batsL <-  dat %>% select(., Plant_Species, fruit_diameter, fruit_length) %>% unique() %>% left_join(batsL, ., by="Plant_Species")
batsL <- bats %>% select(., Frugivore_Species, Plant_Species) %>% mutate(., real=1) %>% left_join(batsL, ., by=c("Frugivore_Species", "Plant_Species"))
batsL$real[is.na(batsL$real)==TRUE] <- 0

batsL %<>% filter(., Frugivore_Species != "Carollia castanea") #wrong gape size

#Sample our overall data
batsL_sample <- batsL[sample(nrow(batsL)/2, replace=T),]



p <- batsL_sample[batsL_sample$real ==1,3:(ncol(batsL_sample)-1)]
bgrd <- batsL_sample[,3:(ncol(batsL_sample)-1)]

mean.p <- colMeans(p, na.rm = TRUE)         # estimated mean of presence points
mean.bgrd <- colMeans(bgrd)   # estimated mean of background points
sigma.p <- stats::cov(p)             # estimated covariance of presence points
sigma.bgrd <- stats::cov(bgrd)

batsL[,3:(ncol(batsL_sample)-1)]

require(np)
f1 <- npudensbw(p, bwmethod='normal-reference')
f0 <- npudensbw(bgrd, bwmethod='normal-reference')
  
f1_fit <- fitted(npudens(bws=f1, edat=batsL[,3:(ncol(batsL_sample)-1)], bwmethod='normal-reference'))
f0_fit<- fitted(npudens(bws=f0, edat=batsL[,3:(ncol(batsL_sample)-1)], bwmethod='normal-reference'))

S <- f1_fit/f0_fit

batsL$S <- S

AUC_PP_Traits_Bats <- roc(data=batsL, response=real, predictor=S)
plot(AUC_PP_Traits_Bats)
AUC_PP_Bats
```

```{r}
#Make long form of just bat intearactions.
birds <- dplyr::filter(dat, Frug_Class=="Aves")
birdsL <- tidyr::expand(birds, Frugivore_Species, Plant_Species)

#Left_Join in our trait data and denote real entries
birdsL <-  birds %>% select(., Frugivore_Species, Frug_Mean_Gape_Size, Frug_Body_Mass) %>% unique() %>% left_join(birdsL, ., by="Frugivore_Species")
birdsL <-  dat %>% select(., Plant_Species, fruit_diameter, fruit_length) %>% unique() %>% left_join(birdsL, ., by="Plant_Species")
birdsL <- birds %>% select(., Frugivore_Species, Plant_Species) %>% mutate(., real=1) %>% left_join(birdsL, ., by=c("Frugivore_Species", "Plant_Species"))
birdsL$real[is.na(birdsL$real)==TRUE] <- 0

birdsL %<>% filter(., Frugivore_Species != "Carollia castanea") #wrong gape size

#Sample our overall data
birdsL_sample <- birdsL[sample(nrow(birdsL)/2, replace=T),]



p <- birdsL_sample[birdsL_sample$real ==1,3:(ncol(birdsL_sample)-1)]
bgrd <- birdsL_sample[,3:(ncol(birdsL_sample)-1)]

mean.p <- colMeans(p, na.rm = TRUE)         # estimated mean of presence points
mean.bgrd <- colMeans(bgrd)   # estimated mean of background points
sigma.p <- stats::cov(p)             # estimated covariance of presence points
sigma.bgrd <- stats::cov(bgrd)

birdsL[,3:(ncol(birdsL_sample)-1)]

require(np)
f1 <- npudensbw(p, bwmethod='normal-reference')
f0 <- npudensbw(bgrd, bwmethod='normal-reference')
  
f1_fit <- fitted(npudens(bws=f1, edat=birdsL[,3:(ncol(birdsL_sample)-1)], bwmethod='normal-reference'))
f0_fit <- fitted(npudens(bws=f0, edat=birdsL[,3:(ncol(birdsL_sample)-1)], bwmethod='normal-reference'))

S <- f1_fit/f0_fit


birdsL$S <- S
birdsL <- birdsL[is.na(birdsL$S)==FALSE,]
birdsL$S[birdsL$S==Inf] <- max(birdsL$S[birdsL$S!=Inf], na.rm=TRUE)+1

AUC_PP_Traits_birds <- roc(data=birdsL, response=real, predictor=S)
plot(AUC_PP_Traits_birds)
AUC_PP_Traits_birds
```

Alright; now that plug and play works for host and plant traits let's add in phylogeny. 

I set all branch lengths to 1 and then force it back to being ultrametric. This feels like bad juju.
```{r}
library(PVR)
library(ape)
library(phytools)

plant_tree <- ape::compute.brlen(plant_tree, 1) #set branch lengths to 1
plant_tree <- phytools::force.ultrametric(plant_tree) #force back to being ultrametric (may be bad)
plant_litter <- PVR::PVRdecomp(phy=plant_tree, type="newick", scale=TRUE)

round((plant_litter@Eigen$values)/sum((plant_litter@Eigen$values)),3)*100 #first three vectors only contain about 28% of variation :(

plant_PhyEig <-data.frame(plant_tree$tip.label, plant_litter@Eigen$vectors[,1:3])

plant_PhyEig %<>% rename(., "Plant_Species"=plant_tree.tip.label)
plant_PhyEig$Plant_Species <- gsub("_ott.*","", x=plant_PhyEig$Plant_Species)
plant_PhyEig$Plant_Species <- gsub(pattern="_", replace=" ", x=plant_PhyEig$Plant_Species)

dat <- left_join(dat, plant_PhyEig, by="Plant_Species")

frug_tree <- ape::compute.brlen(frug_tree, 1) #set branch lengths to 1
frug_tree <- phytools::force.ultrametric(frug_tree) #force back to being ultrametric (may be bad)
frug_litter <- PVR::PVRdecomp(phy=frug_tree, type="newick", scale=TRUE)

round((frug_litter@Eigen$values)/sum((frug_litter@Eigen$values)),3)*100 #first three vectors contain about 33.8% of variation

frug_PhyEig <-data.frame(frug_tree$tip.label, frug_litter@Eigen$vectors[,1:3])

frug_PhyEig %<>% rename(., "Frugivore_Species"=frug_tree.tip.label)
frug_PhyEig$Frugivore_Species <- gsub("_ott.*","", x=frug_PhyEig$Frugivore_Species)
frug_PhyEig$Frugivore_Species <- gsub(pattern="_", replace=" ", x=frug_PhyEig$Frugivore_Species)
colnames(frug_PhyEig) <- c("Frugivore_Species", "fc1", "fc2", "fc3")

dat <- left_join(dat, frug_PhyEig, by="Frugivore_Species")
```



```{r}
#Make long form of just bat intearactions.
bats <- dplyr::filter(dat, Frug_Group=="Bats")
batsL <- tidyr::expand(bats, Frugivore_Species, Plant_Species)

#Left_Join in our trait data and denote real entries
batsL <-  bats %>% select(., Frugivore_Species, fc1, fc2, fc3) %>% unique() %>% left_join(batsL, ., by="Frugivore_Species")
batsL <-  dat %>% select(., Plant_Species, c1, c2, c3) %>% unique() %>% left_join(batsL, ., by="Plant_Species")
batsL <- bats %>% select(., Frugivore_Species, Plant_Species) %>% mutate(., real=1) %>% left_join(batsL, ., by=c("Frugivore_Species", "Plant_Species"))

batsL$real[is.na(batsL$real)==TRUE] <- 0

batsL %<>% filter(., Frugivore_Species != "Carollia castanea") #wrong gape size
batsL %<>% filter(., is.na(c1)==FALSE)
batsL %<>% filter(., is.na(fc1)==FALSE)

#Sample our overall data
batsL_sample <- batsL[sample(nrow(batsL)/2, replace=T),]


p <- batsL_sample[batsL_sample$real ==1,3:(ncol(batsL_sample)-1)]
bgrd <- batsL_sample[,3:(ncol(batsL_sample)-1)]

mean.p <- colMeans(p, na.rm = TRUE)         # estimated mean of presence points
mean.bgrd <- colMeans(bgrd)   # estimated mean of background points
sigma.p <- stats::cov(p)             # estimated covariance of presence points
sigma.bgrd <- stats::cov(bgrd)


require(np)
f1 <- npudensbw(p, bwmethod='normal-reference')
f0 <- npudensbw(bgrd, bwmethod='normal-reference')
  
f1_fit <- fitted(npudens(bws=f1, edat=batsL[,3:(ncol(batsL_sample)-1)], bwmethod='normal-reference'))
f0_fit<- fitted(npudens(bws=f0, edat=batsL[,3:(ncol(batsL_sample)-1)], bwmethod='normal-reference'))

S <- f1_fit/f0_fit

batsL$S <- S

AUC_PP_Phylo_Bats <- roc(data=batsL, response=real, predictor=S)
plot(AUC_PP_Phylo_Bats)
AUC_PP_Phylo_Bats
```
```{r}
#Make long form of just bat intearactions.
birds <- dplyr::filter(dat, Frug_Class=="Aves")
birdsL <- tidyr::expand(birds, Frugivore_Species, Plant_Species)

#Left_Join in our trait data and denote real entries
birdsL <-  birds %>% select(., Frugivore_Species, fc1, fc2, fc3) %>% unique() %>% left_join(birdsL, ., by="Frugivore_Species")
birdsL <-  dat %>% select(., Plant_Species, c1, c2, c3) %>% unique() %>% left_join(birdsL, ., by="Plant_Species")
birdsL <- birds %>% select(., Frugivore_Species, Plant_Species) %>% mutate(., real=1) %>% left_join(birdsL, ., by=c("Frugivore_Species", "Plant_Species"))

birdsL$real[is.na(birdsL$real)==TRUE] <- 0

birdsL %<>% filter(., Frugivore_Species != "Carollia castanea") #wrong gape size
birdsL %<>% filter(., is.na(c1)==FALSE)
birdsL %<>% filter(., is.na(fc1)==FALSE)

#Sample our overall data
birdsL_sample <- birdsL[sample(nrow(birdsL)/2, replace=T),]


p <- birdsL_sample[birdsL_sample$real ==1,3:(ncol(birdsL_sample)-1)]
bgrd <- birdsL_sample[,3:(ncol(birdsL_sample)-1)]

mean.p <- colMeans(p, na.rm = TRUE)         # estimated mean of presence points
mean.bgrd <- colMeans(bgrd)   # estimated mean of background points
sigma.p <- stats::cov(p)             # estimated covariance of presence points
sigma.bgrd <- stats::cov(bgrd)


require(np)
f1 <- npudensbw(p, bwmethod='normal-reference')
f0 <- npudensbw(bgrd, bwmethod='normal-reference')
  
f1_fit <- fitted(npudens(bws=f1, edat=birdsL[,3:(ncol(birdsL_sample)-1)], bwmethod='normal-reference'))
f0_fit<- fitted(npudens(bws=f0, edat=birdsL[,3:(ncol(birdsL_sample)-1)], bwmethod='normal-reference'))

S <- f1_fit/f0_fit

birdsL$S <- S

AUC_PP_Phylo_birds <- roc(data=birdsL, response=real, predictor=S)
plot(AUC_PP_Phylo_birds)
AUC_PP_Phylo_birds
```


```{r}
pdf(file="PlugPlayTestAUC.pdf", height = 8.5, width=11)
par(mfrow=c(2,2))
plot(AUC_PP_Traits_Bats, main="Plug & Play Trait AUC; Bats")
mtext("AUC=0.7621", side=1)
plot(AUC_PP_Phylo_Bats, main="Plug & Play PhyloDecomposition AUC; Bats")
mtext("AUC=0.8211", side=1)
plot(AUC_PP_Traits_birds, main="Plug & Play Trait AUC; Birds")
mtext("AUC=0.6782", side=1)
plot(AUC_PP_Phylo_birds, main="Plug & Play PhyloDecomposition AUC; Birds")
mtext("AUC=0.6009", side=1)

#Plants: 28%, animals 33.8% of variation  
```

