---
title: "Frugivory"
author: "Grant Foster"
date: "9/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(magrittr)
library(igraph)
library(pROC)
library(BIEN)
library(np)
library(PVR)
library(ape)
library(phytools)
```


```{r read in data}
dat <- read.csv("Data.nosync/ATLANTIC_frugivory.csv")
dat %<>% filter(., Frugivore_Species != "Carollia castanea") #This bat has an incorrect gape size
```

```{r set up igraph network}

uniq <- dat %>% select(., Frugivore_Species, Plant_Species) %>% unique()
web <- data.frame(
    from = uniq$Frugivore_Species,
    to = uniq$Plant_Species
  ) %>% graph_from_data_frame(directed=FALSE) #Create a graph with all unique plant-frugivore edges

mat <-get.adjacency(web, type="both", sparse = FALSE) #Create adjacency matrix from our newly created web
#isSymmetric.matrix(mat) #we should get a full symmetric adjacency matrix since we added the type="both" argument above; this makes sure

V(web)$type <- bipartite.mapping(web)$type #Assign our bipartite mapping to type

V(web)$color <-  NA #Create empty node-propert for color
V(web)$color[V(web)$type==TRUE] <- "red"
V(web)$color[V(web)$type==FALSE] <- "blue"
```


```{r basic plot of igraph web eval=FALSE}

plot(web, layout=layout_as_bipartite(web), edge.width=0.25, vertex.label=NA, vertex.size=1)


plot_data <- dat %>% mutate(
        to = factor(Plant_Species, levels = unique(c(dat$Plant_Species, dat$Frugivore_Species))),
        from = factor(Frugivore_Species, levels = unique(c(dat$Plant_Species, dat$Frugivore_Species))))

plot_data <- dat %>% mutate(
        to = factor(Plant_Species),
        from = factor(Frugivore_Species))

ggplot(plot_data, aes(x = from, y = to)) +
      geom_raster() +
      theme_bw() +
      # Because we need the x and y axis to display every node,
      # not just the nodes that have connections to each other,
      # make sure that ggplot does not drop unused factor levels
      scale_x_discrete(drop = FALSE) +
      scale_y_discrete(drop = FALSE) +
      theme(
        # Rotate the x-axis lables so they are legible
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        # Force the plot into a square aspect ratio
        aspect.ratio = 1,
        # Hide the legend (optional)
        legend.position = "none")
```

# I need a phylogeny for my species

The frugivore one I get from tree of life is ok, but the plant one kind of sucks-huge polytomies at really pretty high taxonomic levels. 

Let's get a better plant phylogeny than the one we're working with. 

```{r trimming overall BIEN plant trees, eval=FALSE}
plant_list <- unique(dat$Plant_Species)
plant_list <- gsub(" ","_", plant_list)

BIEN_tree <- BIEN_phylogeny_complete(n_phylogenies = 1, seed=1, replicates = 1)
BIEN_tree$tip.label

BIEN_matches <- plant_list[plant_list %in% BIEN_tree$tip.label==TRUE]

BIEN_subtree <- ape::keep.tip(BIEN_tree, tip=BIEN_matches)
pdf(file="BIEN_planttree.pdf")
plot(BIEN_subtree, show.tip.label=FALSE)
```

```{r load Bien subtrees, eval=TRUE}
load("BIEN_subtree.Rda")
```

Now let's get a good frugivore phylogeny
```{r selecting frugivores to query VertLife, eval=FALSE}
Mammal_List <- filter(dat, Frug_Class=="Mammalia") %>% select(., Frugivore_Species) %>% unique()
Bird_List <- filter(dat, Frug_Class=="Aves") %>% select(., Frugivore_Species) %>% unique()

library(clipr)
clipr::write_clip(Mammal_List[,1]) #This way we can copy our species into the VertLife website to request a tree
clipr::write_clip(Bird_List[,1])
```

```{r read and match vertTrees, eval=TRUE}
Mammal_trees <- ape::read.nexus(file="Data.nosync/VertLife_FrugMam/output.nex")

plot(Mammal_trees[1], show.tip.label = FALSE)

Mammal_List <- gsub(" ", "_", Mammal_List[,1])
Mammal_List[Mammal_List %in% Mammal_trees$tree_1642$tip.label==FALSE] #mammals not in our phylogeny

Bird_trees <- ape::read.nexus(file="Data.nosync/VertLife_FrugBird/output.nex")

plot(Bird_trees[1], show.tip.label = FALSE)

Bird_List <- gsub(" ", "_", Bird_List[,1])
Bird_List[Bird_List %in% Bird_trees$tree_2110$tip.label==FALSE] #mammals not in our phylogeny
```


```{r plot our trees, eval=FALSE}
Mammal_trees <- ape::read.nexus(file="Data.nosync/VertLife_FrugMam/output.nex")
Bird_trees <- ape::read.nexus(file="Data.nosync/VertLife_FrugBird/output.nex")
#pdf(file="Phylogenies.pdf", height=11, width=8.5)

par(mfrow = c(3, 2))
ape::plot.phylo(BIEN_subtree, show.tip.label = FALSE, type="fan", show.node.label = FALSE, main="Plants")
plot(BIEN_subtree, show.tip.label=FALSE)

ape::plot.phylo(Mammal_trees$tree_1642, show.tip.label = FALSE, type="fan", show.node.label = FALSE, main="Mammals")
plot(Mammal_trees$tree_1642, show.tip.label=FALSE)

ape::plot.phylo(Bird_trees$tree_2110, show.tip.label = FALSE, type="fan", show.node.label = FALSE, main="Birds")
plot(Bird_trees$tree_2110, show.tip.label=FALSE)
```


Subsampling our graph into training and test set. Deleting edges from our original graph lets us preserve our full adjacency matrix-if we take out entries from our dataframe instead we have problems later on in about predicting out of bounds. 

```{r graph subsampling workflow, eval=FALSE}
set.seed(24601)

test_web <- igraph::delete.edges(web, sample(1:length(E(web)), 3658, replace = FALSE)) #remove 50% of our data
#plot(web, layout=layout_as_bipartite(web))
#plot(test_web, layout=layout_as_bipartite(test_web))

test_adj <- get.adjacency(test_web, type="both", sparse = FALSE)

#lattice::levelplot(test_adj)
isSymmetric.matrix(test_adj)
```

In order to get it working with plug and play, we need to turn our categorical variables into binary ones
```{r factors to binary}
dat <- dplyr::filter_at(dat, vars(Plant_Form, Frugivory_score, Lipid_Score, Fruit_color, Plant_origin), all_vars(!is.na(.))) #make sure we have scores for all our binary variables
dat %<>% dplyr::filter(., Fruit_color !="") #This Remove blank fruit color entries
dat$Plant_origin[dat$Plant_origin !="native"] <- "nonnative" #condense all nonnative plants into one (can revisit later)

#dat$real <- 1 # note that these are real interactions; important when we expand later

dat_old <- dat
dat$Frugivory_score <- paste("FrugScore", dat$Frugivory_score, sep="") #Make value names unambiguous as column names
dat$Lipid_Score <- paste("LipScore", dat$Lipid_Score, sep="")


categories <- list(Forms=unique(dat$Plant_Form), #take the unique entries of our categorical data and save them in a list; (less typing to assign column names based on them below)
                   Frugivory_score=unique(dat$Frugivory_score),
                   Lipid_Score=unique(dat$Lipid_Score),
                   Fruit_color=unique(dat$Fruit_color),
                   Plant_origin=unique(dat$Plant_origin)
                   )
dat_old <- dat

dat <- dat %>% mutate(bin=1) %>% pivot_wider(., names_from = Plant_Form, values_from = bin)
dat <- dat %>% mutate(bin=1) %>% pivot_wider(., names_from = Frugivory_score, values_from = bin)
dat <- dat %>% mutate(bin=1) %>% pivot_wider(., names_from = Lipid_Score, values_from = bin)
dat <- dat %>% mutate(bin=1) %>% pivot_wider(., names_from = Fruit_color, values_from = bin)
dat <- dat %>% mutate(bin=1) %>% pivot_wider(., names_from = Plant_origin, values_from = bin)

dat[, 32:ncol(dat)][is.na(dat[,32:ncol(dat)])==TRUE] <- 0 #Replae all NA's in our newly created columns with 0

dat <- left_join(dat, dat_old) #add back in our factor columns just in case we want them later
```



#Doing Phylogenetic Decompositions and Adding them to dat

Since we have one tree for the plant species, I just do a singular decomposition and assign the values to the big dataframe. 
```{r add plant phylogeny to dat}
plant_litter <- PVR::PVRdecomp(phy=BIEN_subtree, type="newick", scale=TRUE)

sum((round((plant_litter@Eigen$values)/sum((plant_litter@Eigen$values)),3)*100)[1:5]) #first three vectors only contain about 28% of variation :(

plant_PhyEig <-data.frame(BIEN_subtree$tip.label, plant_litter@Eigen$vectors[,1:5])
colnames(plant_PhyEig) <- c("Plant_Species", paste(colnames(plant_PhyEig)[2:ncol(plant_PhyEig)], "PlDecomp", sep=""))

plant_PhyEig$Plant_Species <- gsub(pattern="_", replace=" ", x=plant_PhyEig$Plant_Species)

dat <- left_join(dat, plant_PhyEig, by="Plant_Species")
```

Frugivore trees-I'm doing this dumb
```{r}
frug_tree <- ape::compute.brlen(frug_tree, 1) #set branch lengths to 1
frug_tree <- phytools::force.ultrametric(frug_tree) #force back to being ultrametric (may be bad)
frug_litter <- PVR::PVRdecomp(phy=frug_tree, type="newick", scale=TRUE)

round((frug_litter@Eigen$values)/sum((frug_litter@Eigen$values)),3)*100 #first three vectors contain about 33.8% of variation

frug_PhyEig <-data.frame(frug_tree$tip.label, frug_litter@Eigen$vectors[,1:3])

frug_PhyEig %<>% rename(., "Frugivore_Species"=frug_tree.tip.label)
frug_PhyEig$Frugivore_Species <- gsub("_ott.*","", x=frug_PhyEig$Frugivore_Species)
frug_PhyEig$Frugivore_Species <- gsub(pattern="_", replace=" ", x=frug_PhyEig$Frugivore_Species)
colnames(frug_PhyEig) <- c("Frugivore_Species", "fc1", "fc2", "fc3")

dat <- left_join(dat, frug_PhyEig, by="Frugivore_Species")
```


Functionalizing Plug n Play
```{r functionalizing PlugNPlay}

#' Plug and Play algorithm. 
#' Applies the plug and play algorithm to data on environmental covariates at presence and background using multivariate Gaussian distribution for f and f1
#'
#'  @param dat input dataframe of plant-frugivore interactions in the format of the atlantic forest csv
#'  @param FrugTraits Names of columns in dat containing Frugivore traits you'd like to classify with. These must be numeric, with factor entries being converted into a series of binary variables. 
#'  @param PlantTraits  Names of columns in dat containing Plant traits you'd like to classify with; same conditions as above
#'  
#' @returns A list containing: 
#'  f1 : Estimated Kernal Density of positive training cases
#'  f0 : Estimated Kernal Density of background training cases
#'  test :  Edgelist of test set; not informed in train
#'  train : The edgelist we trained our model on (randomly sampled 80% of total possible edges)


pp.kde <- function(dat, FrugTraits, PlantTraits){
  #Set up our data into a fully expanded edgelist
  require(tidyr)
  dat$real <- 1 #make a new column denoting all of these edges are real; important when we expand out to all pairwise connections later
  dat <- dplyr::filter_at(dat, vars(c(FrugTraits, PlantTraits)), all_vars(!is.na(.))) #make sure we have data for all our predictors
  full_L <- tidyr::expand(dat, Frugivore_Species, Plant_Species) #Expand to include all possible pairwise plant-frugivore interactions.
  full_real <- dplyr::select(dat, Frugivore_Species, Plant_Species, real) %>% left_join(full_L, ., by = c("Frugivore_Species", "Plant_Species")) #notating which of our edges in this expanded data are real 
  full_real$real[is.na(full_real$real)==TRUE] <- 0

  full_real_frugs <- dat[,c("Frugivore_Species", FrugTraits)] %>% unique() %>% left_join(full_real, ., by="Frugivore_Species") #Add in our frugivore traits

  full_real_both <- dat[,c("Plant_Species", PlantTraits)] %>% unique() %>% left_join(full_real_frugs, ., by="Plant_Species") #add in our plant traits
  
  # Subsample our edgelist
  sample_IDs <- sample(round(nrow(full_real_both)*0.8), replace=F) #subsample out 80% of our edges. 
  test <- full_real_both[-sample_IDs,] #assign 80% train set
  train <- full_real_both[sample_IDs,]

  #set up our data to go into npudens
  p <- train[(train[,'real'] == 1), 4:ncol(train)] #positives cases (p)
  bgrd <- train[,4:ncol(train)] #background (positive and negative) cases

  require(np)
  f1 <- np::npudens(p, bwmethod='normal-reference') 
  f0 <- np::npudens(bgrd, bwmethod='normal-reference')
  model=list(f1, f0, test, train)
  names(model) <- c("f1", "f0", "test", "train")
  #browser
  return(model)
}


predict.pp.kde <- function(model, x){
  require(np)
  f1 <- fitted(npudens(bws = model$f1$bws, edat = x, bwmethod = 'normal-reference'))
  f0 <- fitted(npudens(bws = model$f0$bws, edat = x, bwmethod = 'normal-reference'))
  #browser()
  return(S = f1/f0)
}
```

# Apply Plug and Play to Bats

## Traits - Bats
```{r PnP on Bat Traits}
bats <- dplyr::filter(dat, Frug_Group=="Bats")

bats_kde <- pp.kde(bats, FrugTraits = c("Frug_Body_Mass","Frug_Mean_Gape_Size", "FrugScore1", "FrugScore2", "FrugScore3"), PlantTraits = c("fruit_diameter", "fruit_length", "tree", "palm", "scrub", "yellow", "red", "black", "brown", "green", "LipScore1", "LipScore2"))
#Predictors include continuous variables plus plant form and lipid score

bats_kde$test$S <- predict.pp.kde(model=bats_kde, x=bats_kde$test[,4:ncol(bats_kde$test)]) #Predict on test set and assign suitability scores

AUC_BatsTraits <- roc(data=bats_kde$test, response=real, predictor=S)
plot(AUC_BatsTraits)
AUC_BatsTraits
```
  
## Phylogeny - Bats
  
```{r Setup bat phylogeny}
frug_litter <- PVR::PVRdecomp(phy=Mammal_trees[[1]], type="newick", scale=TRUE)

#sum((round((frug_litter@Eigen$values)/sum((frug_litter@Eigen$values)),3)*100)[1:3]) #first three vectors contain 66.4% of variation!!!
frug_PhyEig <-data.frame(Mammal_trees[[1]]$tip.label, frug_litter@Eigen$vectors[,1:4]) #take first three vectors and make them a dataframe
colnames(frug_PhyEig) <- c("Frugivore_Species", "fc1", "fc2", "fc3", "fc4")

frug_PhyEig$Frugivore_Species <- gsub(pattern="_", replace=" ", x=frug_PhyEig$Frugivore_Species) 

bats_temp <- left_join(bats, frug_PhyEig, by="Frugivore_Species") #Make a new object that includes our phylo information
```
  
Run it
```{r PnP on Bat Phylogeny}
batsPhylo_kde <- pp.kde(bats_temp, FrugTraits = c("fc1", "fc2", "fc3", "fc4"), PlantTraits = c("c1PlDecomp", "c2PlDecomp", "c3PlDecomp", "c4PlDecomp"))
#Predictors include continuous variables plus plant form and lipid score

batsPhylo_kde$test$S <- predict.pp.kde(model=batsPhylo_kde, x=batsPhylo_kde$test[,4:ncol(batsPhylo_kde$test)]) #Predict on test set and assign suitability scores

AUC_BatsPhylo <- roc(data=batsPhylo_kde$test, response=real, predictor=S)
plot(AUC_BatsPhylo)
AUC_BatsPhylo
```

  
  

### Apply Plug and Play to Birds

## Traits - Birds
This takes a bit (~5min or so on my machine)
```{r PnP on Bird Traits}
birds <- dplyr::filter(dat, Frug_Class=="Aves")

birds_kde <- pp.kde(birds, FrugTraits = c("Frug_Body_Mass","Frug_Mean_Gape_Size", "FrugScore1", "FrugScore2", "FrugScore3"), PlantTraits = c("fruit_diameter", "fruit_length", "tree","liana", "palm", "scrub", "yellow", "red", "black", "brown", "green", "LipScore1", "LipScore2", "LipScore3"))#Predictors include continuous variables plus plant form and lipid score

birds_kde$test$S <- predict.pp.kde(model=birds_kde, x=birds_kde$test[,4:ncol(birds_kde$test)]) #Predict on test set and assign suitability scores

AUC_birdsTraits <- roc(data=birds_kde$test, response=real, predictor=S)
plot(AUC_birdsTraits)
AUC_birdsTraits
```
## Phylogeny - Birds
  
```{r Setup Bird phylogeny}
frug_litter <- PVR::PVRdecomp(phy=Bird_trees[[5]], type="newick", scale=TRUE)

sum((round((frug_litter@Eigen$values)/sum((frug_litter@Eigen$values)),3)*100)[1:3]) #first three vectors contain 47% of variation
frug_PhyEig <-data.frame(Bird_trees[[5]]$tip.label, frug_litter@Eigen$vectors[,1:3]) #take first three vectors and make them a dataframe
colnames(frug_PhyEig) <- c("Frugivore_Species", "fc1", "fc2", "fc3")

frug_PhyEig$Frugivore_Species <- gsub(pattern="_", replace=" ", x=frug_PhyEig$Frugivore_Species) 

birds_temp <- left_join(birds, frug_PhyEig, by="Frugivore_Species") #Make a new object that includes our phylo information
```
  
Run it
```{r PnP on Bird Phylogeny}
birdsPhylo_kde <- pp.kde(birds_temp, FrugTraits = c("fc1", "fc2", "fc3"), PlantTraits = c("c1PlDecomp", "c2PlDecomp", "c3PlDecomp", "c4PlDecomp"))
#Predictors include continuous variables plus plant form and lipid score

birdsPhylo_kde$test$S <- predict.pp.kde(model=birdsPhylo_kde, x=birdsPhylo_kde$test[,4:ncol(birdsPhylo_kde$test)]) #Predict on test set and assign suitability scores

AUC_BirdsPhylo <- roc(data=birdsPhylo_kde$test, response=real, predictor=S)
plot(AUC_BirdsPhylo)
AUC_BirdsPhylo
```
That looks terrible. Why does it?

```{r plot our results so far}
#pdf(file="PlugPlayTestAUC.pdf", height = 8.5, width=11)
par(mfrow=c(2,2))
plot(AUC_BatsTraits, main="Plug & Play Trait AUC; Bats")
mtext(paste("AUC", round(AUC_BatsTraits$auc, 3), sep="="), side=1)
plot(AUC_BatsPhylo, main="Plug & Play PhyloDecomposition AUC; Bats")
mtext(paste("AUC", round(AUC_BatsPhylo$auc, 3), sep="="), side=1)
plot(AUC_birdsTraits, main="Plug & Play Trait AUC; Birds")
mtext(paste("AUC", round(AUC_birdsTraits$auc, 3), sep="="), side=1)
plot(AUC_BirdsPhylo, main="Plug & Play PhyloDecomposition AUC; Birds")
mtext(paste("AUC", round(AUC_BirdsPhylo$auc, 3), sep="="), side=1)
```
Latent Link Prediction Methods

Tad's Method
  locSim <- rowMeans(as.matrix(vegan::vegdist(intMat)))	
  pathSim <- rowMeans(as.matrix(vegan::vegdist(t(intMat))))
  simMat <- locSim %*% t(pathSim)	
```{r}

mat_assym <- as.data.frame.matrix(table(dat$Frugivore_Species, dat$Plant_Species)) #Rows = Frugivores, Columns = Plants
frugSim <- rowMeans(as.matrix(vegan::vegdist(mat_assym)))	
plantSim <- rowMeans(as.matrix(vegan::vegdist(t(mat_assym))))
simMat <- frugSim %*% t(plantSim)	

data.frame(host = rownames(mat_assym), frugSim)
df$similarity <- as.vector(simMat)
```

```{r}
  bats$real <- 1 #make a new column denoting all of these edges are real; important when we expand out to all pairwise connections later
  full_L <- tidyr::expand(bats, Frugivore_Species, Plant_Species) #Expand to include all possible pairwise plant-frugivore interactions.
  full_real <- dplyr::select(bats, Frugivore_Species, Plant_Species, real) %>% left_join(full_L, ., by = c("Frugivore_Species", "Plant_Species")) #notating which of our edges in this expanded data are real 
  full_real$real[is.na(full_real$real)==TRUE] <- 0
  full_real %<>% unique(.)
  
  mat_assym <- pivot_wider(full_real, names_from = Plant_Species, values_from = real)
  mat_assym %<>% column_to_rownames(., var="Frugivore_Species")
  
  frugSim <- rowMeans(as.matrix(vegan::vegdist(mat_assym)))	
  plantSim <- rowMeans(as.matrix(vegan::vegdist(t(mat_assym))))
  simMat <- frugSim %*% t(plantSim)	

  full_real$similarity <- as.vector(simMat)

  # Subsample our edgelist
  sample_IDs <- sample(round(nrow(full_real)*0.8), replace=F) #subsample out 80% of our edges. 
  test <- full_real[-sample_IDs,] #assign 80% train set
  train <- full_real[sample_IDs,]

  #set up our data to go into npudens
  p <- train[(train[,'real'] == 1), 4:ncol(train)] #positives cases (p)
  bgrd <- train[,4:ncol(train)] #background (positive and negative) cases

  require(np)
  f1 <- np::npudens(p, bwmethod='normal-reference') 
  f0 <- np::npudens(bgrd, bwmethod='normal-reference')
  model=list(f1, f0, test, train)
  names(model) <- c("f1", "f0", "test", "train")
  #browser
  
model$test$S <- predict.pp.kde(model=model, x=model$test[,4:ncol(model$test)]) #Predict on test set and assign suitability scores

AUC_BatsLatent <- roc(data=model$test, response=real, predictor=S)
plot(AUC_BatsLatent, main="vegdist latent: Bats")
AUC_BatsLatent
```

