---
title: "Fruglink Full Network Predictions"
author: "Grant Foster"
date: "2/11/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
dat <- birdsimportance

woodedWalk_NoSplit <- function(dat, FrugTraits, PlantTraits, class_balancing=FALSE, balance_ratio=3, output_type="predictions"){
  if((output_type %in% c("rfobject", "predictions"))==FALSE){
    stop("Invalid output_type: choices are performance or rfobject")
  }
  #Set up our data into a fully expanded edgelist
  require(tidyr)
  dat$real <- 1 #make a new column denoting all of these edges are real; important when we expand out to all pairwise connections later
  dat <- dplyr::filter_at(dat, vars(c(FrugTraits, PlantTraits)), all_vars(!is.na(.))) #make sure we have data for all our predictors
  
  dat <- dplyr::select_at(dat, vars(c(Frugivore_Species, Plant_Species, real, FrugTraits, PlantTraits))) #Select our relevant predictors
  dat %<>% unique(.) #Make sure we only have unique entries
  full_L <- tidyr::expand(dat, Frugivore_Species, Plant_Species) #Expand to include all possible pairwise plant-frugivore interactions.
  full_real <- dplyr::select(dat, Frugivore_Species, Plant_Species, real) %>% left_join(full_L, ., by = c("Frugivore_Species", "Plant_Species")) #notating which of our edges in this expanded data are real 
  full_real$real[is.na(full_real$real)==TRUE] <- 0

  full_real_frugs <- dat[,c("Frugivore_Species", FrugTraits)] %>% unique() %>% left_join(full_real, ., by="Frugivore_Species") #Add in our frugivore traits

  full_real_both <- dat[,c("Plant_Species", PlantTraits)] %>% unique() %>% left_join(full_real_frugs, ., by="Plant_Species") #add in our plant traits
  
  full_real_both$real <- as.factor(full_real_both$real)
  
  #if(class_balancing==TRUE){
    # Subsample our edgelist
    #ones <- full_real_both %>% dplyr::filter(., real==1)
    
    #zeroes <- full_real_both %>% dplyr::filter(., real==0)
    #zero_IDs <- sample(x=1:nrow(zeroes), size=nrow(ones)*balance_ratio, replace=F) #Ratio of train to test 1:3
    #keptZeroes <- zeroes[zero_IDs,]
    
    #nrow(rbind(trainZeroes, testZeroes, trainOnes, testOnes)) == nrow(full_real_both) #Making sure we're not losing anything
    #balanced <- rbind(ones, keptZeroes)
    #full_real_both <- balanced
  #}
  
  rf<-randomForest::randomForest(real ~. -Frugivore_Species - Plant_Species, data=full_real_both, ntree=100)
  #ROC <- roc(full_real_both$real, rf$votes[,2])
  predictions <- stats::predict(rf, newdata=full_real_both, type="prob")
  
  if(output_type=="predictions"){
    output <- full_real_both %>% select(., "Frugivore_Species", "Plant_Species", real)
    output$S <- predictions[,2]
    return(output)
  }
  
  if(output_type=="rfobject"){
    output <- rf
  return(output)
  }
  #return(output)
}
```


```{r}
  frug_litter <- PVR::PVRdecomp(phy=Bird_trees[[round(1)]], type="newick", scale=TRUE)
  frug_PhyEig <-data.frame(Bird_trees[[5]]$tip.label, frug_litter@Eigen$vectors[,1:3]) #take first three vectors and make them a dataframe
  colnames(frug_PhyEig) <- c("Frugivore_Species", "fc1", "fc2", "fc3")
  frug_PhyEig$Frugivore_Species <- gsub(pattern="_", replace=" ", x=frug_PhyEig$Frugivore_Species) 
  birds <- left_join(birds, frug_PhyEig, by="Frugivore_Species") #Make a new object that includes our phylo information


p_phylo <- c("c1PlDecomp", "c2PlDecomp", "c3PlDecomp", "c4PlDecomp")
p_traits <- c("fruit_diameter", "fruit_length", "tree","liana", "palm", "scrub", "yellow", "red", "black", "brown", "green", "Lipid_Score")
#p_latent <- c("plantSim", "Psvd1", "Psvd2", "Psvd3") #Both methods together
p_latent <- c("Psvd1", "Psvd2", "Psvd3")

f_phylo <- c("fc1", "fc2", "fc3")
f_traits <- c("Frug_Body_Mass","Frug_Mean_Gape_Size", "Frugivory_score")
#f_latent <- c("frugSim", "Fsvd1", "Fsvd2", "Fsvd3")
f_latent <- c("Fsvd1", "Fsvd2", "Fsvd3")
```


```{r}
Phy <- woodedWalk_NoSplit(birds, FrugTraits = c(f_phylo), PlantTraits = c(p_phylo), class_balancing = FALSE, output_type = "predictions")
colnames(Phy) <- c("Frugivore_Species", "Plant_Species", "real", "Phy")

Traits <- woodedWalk_NoSplit(birds, FrugTraits = c(f_traits), PlantTraits = c(p_traits), class_balancing = FALSE, output_type = "predictions")
colnames(Traits) <- c("Frugivore_Species", "Plant_Species", "real", "Traits")

Latent <- woodedWalk_NoSplit(birds, FrugTraits = c(f_latent), PlantTraits = c(p_latent), class_balancing = FALSE, output_type = "predictions")
colnames(Latent) <- c("Frugivore_Species", "Plant_Species", "real", "Latent")


PhyTraits <- woodedWalk_NoSplit(birds, FrugTraits = c(f_phylo, f_traits), PlantTraits = c(p_phylo, p_traits), class_balancing = FALSE, output_type = "predictions")
colnames(PhyTraits) <- c("Frugivore_Species", "Plant_Species", "real", "PhyTraits")

TraitsLatent <- woodedWalk_NoSplit(birds, FrugTraits = c(f_latent, f_traits), PlantTraits = c(p_latent, p_traits), class_balancing = FALSE, output_type = "predictions")
colnames(TraitsLatent) <- c("Frugivore_Species", "Plant_Species", "real", "TraitsLatent")


PhyLatent <- woodedWalk_NoSplit(birds, FrugTraits = c(f_latent, f_phylo), PlantTraits = c(p_latent, p_phylo), class_balancing = FALSE, output_type = "predictions")
colnames(PhyLatent) <- c("Frugivore_Species", "Plant_Species", "real", "PhyLatent")


Trio <- woodedWalk_NoSplit(birds, FrugTraits = c(f_phylo, f_latent, f_traits), PlantTraits = c(p_phylo, p_latent, p_traits), class_balancing = FALSE, output_type = "predictions")
colnames(Trio) <- c("Frugivore_Species", "Plant_Species", "real", "Trio")


composite <- Traits %>% select(., -real) %>% left_join(Phy, ., by = c("Frugivore_Species", "Plant_Species"))
composite <- Latent %>% select(., -real) %>% dplyr::left_join(composite, ., by=c("Frugivore_Species", "Plant_Species"))
composite <- PhyTraits %>% select(., -real) %>% left_join(composite, .)
composite <- TraitsLatent %>% select(., -real) %>% left_join(composite, .)
composite <- PhyLatent %>% select(., -real) %>% left_join(composite, .)
composite <- Trio %>% select(., -real) %>% left_join(composite, .)

composite %<>% filter(., is.na(Traits)==FALSE)
composite %<>% filter(., is.na(Phy)==FALSE)
composite %<>% filter(., is.na(Trio)==FALSE)

range(table(composite$Frugivore_Species, composite$Plant_Species))
```


```{r}
cor.matrix <- composite%>% select(., -Frugivore_Species, -Plant_Species, -real) %>%
  as.matrix %>%
  cor %>%
  as.data.frame %>%
  rownames_to_column(var = 'var1') %>%
  gather(var2, value, -var1)

cor.conventional <- composite%>% select(., -Frugivore_Species, -Plant_Species, -real) %>%
  as.matrix %>%
  cor

spear.matrix <- composite%>% select(., -Frugivore_Species, -Plant_Species, -real) %>%
  as.matrix %>%
  cor(method="spearman") %>%
  as.data.frame %>%
  rownames_to_column(var = 'var1') %>%
  gather(var2, value, -var1)

spear.conventional <- composite%>% select(., -Frugivore_Species, -Plant_Species, -real) %>%
  as.matrix %>%
  cor(method="spearman")
```


```{r}
predicts <- dplyr::filter(composite, real==0)

binQuants <- function(input, quantil=0.95){
  input[input > quantile(input , quantil)] <- 1
  input[input<1] <- 0
  return(input)
}

predicts$Phy <- binQuants(predicts$Phy)
predicts$Traits <- binQuants(predicts$Traits)
predicts$Latent <- binQuants(predicts$Latent)
predicts$PhyTraits <- binQuants(predicts$PhyTraits)
predicts$PhyLatent <- binQuants(predicts$PhyLatent)
predicts$TraitsLatent <- binQuants(predicts$TraitsLatent)
predicts$Trio <- binQuants(predicts$Trio)


predicts <- mutate(predicts, sum_agree=Phy+Traits+Latent+PhyTraits+TraitsLatent+PhyLatent+Trio)
predicts %<>% filter(., sum_agree > 0)
table(predicts$sum_agree)

predicts <- mutate(predicts, ordering=Latent*4+
                     Phy*8+
                     PhyLatent*16+
                     PhyTraits*32+
                     TraitsLatent*64+
                     Traits*128+
                     Trio*256)


predicts_best <- dplyr::filter(predicts, sum_agree > 4)
predicts_best <- predicts
predicts_best$Link <- paste(predicts_best$Frugivore_Species, predicts_best$Plant_Species, sep="-")

predicts_best[,4:10][predicts_best[,4:10]==0] <- NA

predicts_best$Link <- as.factor(predicts_best$Link)
predicts_best$Link <- reorder(predicts_best$Link, predicts_best$ordering)


predicts_best$Link <- reorder(predicts_best$Link, predicts_best$sum_agree)
predicts_best$Link <- reorder(predicts_best$Link, -predicts_best$Trio)
predicts_best$Link <- reorder(predicts_best$Link, -predicts_best$Phy)
predicts_best$Link <- reorder(predicts_best$Link, -predicts_best$Traits)
predicts_best$Link <- reorder(predicts_best$Link, -predicts_best$PhyTraits)
predicts_best$Link <- reorder(predicts_best$Link, -predicts_best$Latent)
predicts_best$Link <- reorder(predicts_best$Link, -predicts_best$PhyLatent)
predicts_best$Link <- reorder(predicts_best$Link, -predicts_best$Trio)


ggplot(data=predicts_best, aes(y=Link, x=as.factor("Phy"), fill=Phy))+geom_tile()+
geom_tile(aes(y=Link, x=as.factor("Latent"), fill=Latent))+
  geom_tile(aes(y=Link, x=as.factor("Traits"), fill=Traits))+
  geom_tile(aes(y=Link, x=as.factor("PhyTraits"), fill=PhyTraits))+
  geom_tile(aes(y=Link, x=as.factor("PhyLatent"), fill=PhyLatent))+
  geom_tile(aes(y=Link, x=as.factor("TraitLatent"), fill=TraitsLatent))+
  geom_tile(aes(y=Link, x=as.factor("Trio"), fill=Trio))+
  theme(axis.text.y = element_blank())
```


```{r}
table(predicts$sum_agree)

loneHater <- predicts[predicts$sum_agree==6,]


temp <- c(table((loneHater$Phy==0))["TRUE"], "Phy")
temp <- rbind(temp, c(table((loneHater$Traits==0))["TRUE"], "Traits"))
temp <- rbind(temp, c(table((loneHater$Latent==0))["TRUE"], "Latent"))
temp <- rbind(temp, c(table((loneHater$PhyTraits==0))["TRUE"], "PhyTraits"))
temp <- rbind(temp, c(table((loneHater$TraitsLatent==0))["TRUE"], "TraitsLatent"))
temp <- rbind(temp, c(table((loneHater$PhyLatent==0))["TRUE"], "PhyLatent"))
temp <- rbind(temp, c(table((loneHater$Trio==0))["TRUE"], "Trio"))

dim(temp)

temp <- data.frame(temp)
rownames(temp) <- NULL
colnames(temp) <- c("Single_Hater_Count", "Model")
temp$Single_Hater_Count <- as.numeric(temp$Single_Hater_Count)



test <- dplyr::filter(predicts, sum_agree==5)
nombres <- colnames(test)[4:10]
doubleHaters <- matrix(nrow = length(nombres), ncol = length(nombres))
colnames(doubleHaters) <- nombres
rownames(doubleHaters) <- nombres
for (i in 4:10){
  for (j in 4:10){
    x <- test[,i]+test[,j]
    doubleHaters[i-3,j-3] <- as.numeric(table(x)["0"])
  }
}

doubleHaters[is.na(doubleHaters)==TRUE] <- 0
diag(doubleHaters)<-NA

sum(temp$Single_Hater_Count)
sum(doubleHaters, na.rm = TRUE)/2

colnames(cor.matrix) <- c("Model_1", "Model_2", "correlation")
colnames(spear.matrix) <- c("Model_1", "Model_2", "correlation")

Tab1 <- kableExtra::kable(temp, caption = "Table 1: Number of times when a given model was the only one to disagree on a link's placement in the top 95%. N=225") %>% kableExtra::kable_classic()
Tab2 <- kableExtra::kable(doubleHaters, caption = "Table 2: Number of times when any 2 models disagreed with the other 5 on a link's placement in the top 95%. N=272") %>% kableExtra::kable_styling()
Tab3 <- kableExtra::kable(round(cor.conventional, 2), caption = "Table 3:Pairwise pearson corelations of suitibility values") %>% kableExtra::kable_styling()
Tab4 <- kableExtra::kable(round(spear.conventional, 2), caption = "Table 4: Pairwise Spearman's corelations of suitibility values") %>% kableExtra::kable_styling()


#kableExtra::save_kable(Tab1, file="Tables/Tab1.pdf")
#kableExtra::save_kable(Tab2, file="Tables/Tab2.pdf")
#kableExtra::save_kable(Tab3, file="Tables/Tab3.pdf")
#kableExtra::save_kable(Tab4, file="Tables/Tab4.pdf")
```

```{r}
spear.matrix_full <- cor(composite[,4:10], method="spearman")
corrplot::corrplot(spear.matrix_full, type = "lower", diag=T, order = "AOE", main="Spearman: all")

corr.matrix_full <- cor(composite[,4:10], method="pearson")
corrplot::corrplot(corr.matrix_full, type = "lower", diag=T, order = "AOE", main="Pearson: all")

spear.matrix_unreal <- cor(composite_unreal[,4:10], method="spearman")
corrplot::corrplot(spear.matrix_unreal, type = "lower", diag=T, order = "AOE", main="Spearman: unobserved")

corr.matrix_unreal <- cor(composite_unreal[,4:10], method="pearson")
corrplot::corrplot(corr.matrix_unreal, type = "lower", diag=T, order = "AOE", main="Pearson: unobserved")

pdf(file="ESA_Figs/CorrPlots.pdf", width=8.5, height=11)

par(mfrow = c(2, 2), oma=c(0,0,2,0))
corrplot::corrplot(spear.matrix_full, type = "lower", diag=T, main="Spearman: All Links", mar=c(0,0,2,0), method="ellipse")
corrplot::corrplot(corr.matrix_full, type = "lower", diag=T, main="Pearson: All Links", mar=c(0,0,2,0), method="ellipse")
corrplot::corrplot(spear.matrix_unreal, type = "lower", diag=T, main="Spearman: Unobserved Links", mar=c(0,0,2,0), method="ellipse")
corrplot::corrplot(corr.matrix_unreal, type = "lower", diag=T, main="Pearson: Unobserved Links", mar=c(0,0,2,0), method="ellipse")

par(mfrow = c(2, 2), oma=c(0,0,2,0))
corrplot::corrplot.mixed(spear.matrix_full, lower="number", upper="ellipse", diag=T, main="Spearman: All Links", mar=c(0,0,2,0))
corrplot::corrplot(corr.matrix_full, type = "lower", diag=T, main="Pearson: All Links", mar=c(0,0,2,0), method="ellipse")
corrplot::corrplot(spear.matrix_unreal, type = "lower", diag=T, main="Spearman: Unobserved Links", mar=c(0,0,2,0), method="ellipse")
corrplot::corrplot(corr.matrix_unreal, type = "lower", diag=T, main="Pearson: Unobserved Links", mar=c(0,0,2,0), method="ellipse")

dev.off()


pdf(file="ESA_Figs/MixedCorrPlots.pdf", width=8.5, height=9.2)
par(mfrow = c(2, 2), oma=c(0,0,2,0))
corrplot::corrplot.mixed(spear.matrix_full, 
	lower='number', upper='ellipse',
	tl.col=1, 
	addgrid.col=grey(0.5,0.5), 
	mar=c(0,0,2,0), main="Spearman: All Links", tl.pos = "lt",
	diag = "u"
)

corrplot::corrplot.mixed(corr.matrix_full, 
	lower='number', upper='ellipse',
	tl.col=1, 
	addgrid.col=grey(0.5,0.5), 
	mar=c(0,0,2,0), main="Pearson: All Links", tl.pos = "lt",
	diag = "u"
)
corrplot::corrplot.mixed(spear.matrix_unreal, 
	lower='number', upper='ellipse',
	tl.col=1, 
	addgrid.col=grey(0.5,0.5), 
	mar=c(0,0,2,0), main="Spearman: Unobserved Links", tl.pos = "lt",
	diag = "u"
)

corrplot::corrplot.mixed(corr.matrix_unreal, 
	lower='number', upper='ellipse',
	tl.col=1, 
	addgrid.col=grey(0.5,0.5), 
	mar=c(0,0,2,0), main="Pearson: Unobserved Links", tl.pos = "lt",
	diag = "u"
)
```

Trying a different kind of plot

```{r}
web <- data.frame(
    from = birds$Frugivore_Species,
    to = birds$Plant_Species
  ) %>% graph_from_data_frame(directed=FALSE) #Create a graph with all unique plant-frugivore edges

V(web)$degree

degree <- data.frame(degree=igraph::degree(web)) %>% rownames_to_column(., var="species")

Frug_degree <- dplyr::filter(degree, species %in% composite$Frugivore_Species) %>% dplyr::rename(., Frugdegree=degree, Frugivore_Species=species)
Plant_degree <- dplyr::filter(degree, species %in% composite$Plant_Species) %>% dplyr::rename(., Plantdegree=degree, Plant_Species=species)

composite <- left_join(composite, Frug_degree, by="Frugivore_Species")
composite <- left_join(composite, Plant_degree, by="Plant_Species")
composite$sumdegree <- composite$Plantdegree + composite$Frugdegree
```



```{r}
colnames(composite)
composite_unreal <- dplyr::filter(composite, real==0)
composite_long <- pivot_longer(composite_unreal, cols=Phy:Trio, names_to = "model")


p1 <- ggplot(composite_long, aes(x=sumdegree, y=value, color=model))+geom_point()




degree_correlations <- data.frame(model=c("Latent", "Phy", "Traits", "PhyTraits", "TraitsLatent", "PhyLatent", "Trio"),
  cor=c(cor.test(composite_unreal$Latent, composite_unreal$sumdegree)$estimate,
cor.test(composite_unreal$Phy, composite_unreal$sumdegree)$estimate,
cor.test(composite_unreal$Traits, composite_unreal$sumdegree)$estimate,
cor.test(composite_unreal$PhyTraits, composite_unreal$sumdegree)$estimate,
cor.test(composite_unreal$TraitsLatent, composite_unreal$sumdegree)$estimate,
cor.test(composite_unreal$PhyLatent, composite_unreal$sumdegree)$estimate,
cor.test(composite_unreal$Trio, composite_unreal$sumdegree)$estimate),
low95=c(cor.test(composite_unreal$Latent, composite_unreal$sumdegree)$conf.int[1],
cor.test(composite_unreal$Phy, composite_unreal$sumdegree)$conf.int[1],
cor.test(composite_unreal$Traits, composite_unreal$sumdegree)$conf.int[1],
cor.test(composite_unreal$PhyTraits, composite_unreal$sumdegree)$conf.int[1],
cor.test(composite_unreal$TraitsLatent, composite_unreal$sumdegree)$conf.int[1],
cor.test(composite_unreal$PhyLatent, composite_unreal$sumdegree)$conf.int[1],
cor.test(composite_unreal$Trio, composite_unreal$sumdegree)$conf.int[1]),
high95=c(cor.test(composite_unreal$Latent, composite_unreal$sumdegree)$conf.int[2],
cor.test(composite_unreal$Phy, composite_unreal$sumdegree)$conf.int[2],
cor.test(composite_unreal$Traits, composite_unreal$sumdegree)$conf.int[2],
cor.test(composite_unreal$PhyTraits, composite_unreal$sumdegree)$conf.int[2],
cor.test(composite_unreal$TraitsLatent, composite_unreal$sumdegree)$conf.int[2],
cor.test(composite_unreal$PhyLatent, composite_unreal$sumdegree)$conf.int[2],
cor.test(composite_unreal$Trio, composite_unreal$sumdegree)$conf.int[2]))

plant_degreecorrelations <- data.frame(model=c("Latent", "Phy", "Traits", "PhyTraits", "TraitsLatent", "PhyLatent", "Trio"),
  cor=c(cor.test(composite_unreal$Latent, composite_unreal$Plantdegree)$estimate,
cor.test(composite_unreal$Phy, composite_unreal$Plantdegree)$estimate,
cor.test(composite_unreal$Traits, composite_unreal$Plantdegree)$estimate,
cor.test(composite_unreal$PhyTraits, composite_unreal$Plantdegree)$estimate,
cor.test(composite_unreal$TraitsLatent, composite_unreal$Plantdegree)$estimate,
cor.test(composite_unreal$PhyLatent, composite_unreal$Plantdegree)$estimate,
cor.test(composite_unreal$Trio, composite_unreal$Plantdegree)$estimate),
low95=c(cor.test(composite_unreal$Latent, composite_unreal$Plantdegree)$conf.int[1],
cor.test(composite_unreal$Phy, composite_unreal$Plantdegree)$conf.int[1],
cor.test(composite_unreal$Traits, composite_unreal$Plantdegree)$conf.int[1],
cor.test(composite_unreal$PhyTraits, composite_unreal$Plantdegree)$conf.int[1],
cor.test(composite_unreal$TraitsLatent, composite_unreal$Plantdegree)$conf.int[1],
cor.test(composite_unreal$PhyLatent, composite_unreal$Plantdegree)$conf.int[1],
cor.test(composite_unreal$Trio, composite_unreal$Plantdegree)$conf.int[1]),
high95=c(cor.test(composite_unreal$Latent, composite_unreal$Plantdegree)$conf.int[2],
cor.test(composite_unreal$Phy, composite_unreal$Plantdegree)$conf.int[2],
cor.test(composite_unreal$Traits, composite_unreal$Plantdegree)$conf.int[2],
cor.test(composite_unreal$PhyTraits, composite_unreal$Plantdegree)$conf.int[2],
cor.test(composite_unreal$TraitsLatent, composite_unreal$Plantdegree)$conf.int[2],
cor.test(composite_unreal$PhyLatent, composite_unreal$Plantdegree)$conf.int[2],
cor.test(composite_unreal$Trio, composite_unreal$Plantdegree)$conf.int[2]))

frug_degreecorrelations <- data.frame(model=c("Latent", "Phy", "Traits", "PhyTraits", "TraitsLatent", "PhyLatent", "Trio"),
  cor=c(cor.test(composite_unreal$Latent, composite_unreal$Frugdegree)$estimate,
cor.test(composite_unreal$Phy, composite_unreal$Frugdegree)$estimate,
cor.test(composite_unreal$Traits, composite_unreal$Frugdegree)$estimate,
cor.test(composite_unreal$PhyTraits, composite_unreal$Frugdegree)$estimate,
cor.test(composite_unreal$TraitsLatent, composite_unreal$Frugdegree)$estimate,
cor.test(composite_unreal$PhyLatent, composite_unreal$Frugdegree)$estimate,
cor.test(composite_unreal$Trio, composite_unreal$Frugdegree)$estimate),
low95=c(cor.test(composite_unreal$Latent, composite_unreal$Frugdegree)$conf.int[1],
cor.test(composite_unreal$Phy, composite_unreal$Frugdegree)$conf.int[1],
cor.test(composite_unreal$Traits, composite_unreal$Frugdegree)$conf.int[1],
cor.test(composite_unreal$PhyTraits, composite_unreal$Frugdegree)$conf.int[1],
cor.test(composite_unreal$TraitsLatent, composite_unreal$Frugdegree)$conf.int[1],
cor.test(composite_unreal$PhyLatent, composite_unreal$Frugdegree)$conf.int[1],
cor.test(composite_unreal$Trio, composite_unreal$Frugdegree)$conf.int[1]),
high95=c(cor.test(composite_unreal$Latent, composite_unreal$Frugdegree)$conf.int[2],
cor.test(composite_unreal$Phy, composite_unreal$Frugdegree)$conf.int[2],
cor.test(composite_unreal$Traits, composite_unreal$Frugdegree)$conf.int[2],
cor.test(composite_unreal$PhyTraits, composite_unreal$Frugdegree)$conf.int[2],
cor.test(composite_unreal$TraitsLatent, composite_unreal$Frugdegree)$conf.int[2],
cor.test(composite_unreal$PhyLatent, composite_unreal$Frugdegree)$conf.int[2],
cor.test(composite_unreal$Trio, composite_unreal$Frugdegree)$conf.int[2]))





degree_correlations$model <- as.factor(degree_correlations$model)
degree_correlations$model <- reorder(degree_correlations$model, degree_correlations$cor)
ggplot(data=degree_correlations, aes(x=model, y=cor))+geom_col()+geom_errorbar(aes(ymin=low95, ymax=high95))+ylab("Correlation between Link Suitability and Plant+Frugivore degree")


frug_degreecorrelations$model <- as.factor(frug_degreecorrelations$model)
frug_degreecorrelations$model <- reorder(frug_degreecorrelations$model, frug_degreecorrelations$cor)
ggplot(data=frug_degreecorrelations, aes(x=model, y=cor))+geom_col()+geom_errorbar(aes(ymin=low95, ymax=high95))+ylab("Correlation between Link Suitability and Frugivore degree")

plant_degreecorrelations$model <- as.factor(plant_degreecorrelations$model)
plant_degreecorrelations$model <- reorder(plant_degreecorrelations$model, plant_degreecorrelations$cor)
ggplot(data=plant_degreecorrelations, aes(x=model, y=cor))+geom_col()+geom_errorbar(aes(ymin=low95, ymax=high95))+ylab("Correlation between Link Suitability and Plant degree")


pdf()

p2 <- ggplot(data=degree_correlations, aes(x=model, y=cor))+geom_col()+geom_errorbar(aes(ymin=low95, ymax=high95))+ylab("Correlation between Link Suitability and Plant+Frugivore degree")



pdf(file="degree_v_suitability.pdf", width=11, height=11)
gridExtra::grid.arrange(p1, p2, nrow = 2)

```



Try above plot again, but this time with continuous fill

```{r}

composite <- dplyr::mutate(composite, sum=Phy+Traits+Latent+PhyTraits+TraitsLatent+PhyLatent+Trio)

key <- composite
key$Phy <- binQuants(key$Phy, quantil = 0.9)
key$Traits <- binQuants(key$Traits, quantil = 0.9)
key$Latent <- binQuants(key$Latent, quantil = 0.9)
key$PhyTraits <- binQuants(key$PhyTraits, quantil = 0.9)
key$PhyLatent <- binQuants(key$PhyLatent, quantil = 0.9)
key$TraitsLatent <- binQuants(key$TraitsLatent, quantil = 0.9)
key$Trio <- binQuants(key$Trio, quantil = 0.9)
key$sum_agree <- key$Phy+ key$Traits + key$Latent + key$PhyTraits + key$PhyLatent + key$TraitsLatent + key$Trio


top <- composite[which(key$sum_agree>0),]
```



```{r}
top$Link <- paste(top$Frugivore_Species, top$Plant_Species, sep="-")
selection <- top[sample(1:nrow(top), size=100, replace = FALSE),]

ggplot(data=selection, aes(y=Link, x=as.factor("Phy"), fill=Phy))+
  geom_tile()+
  theme(axis.text.y = element_blank())+
  geom_tile(aes(y=Link, x=as.factor("Latent"), fill=Latent))+
  geom_tile(aes(y=Link, x=as.factor("Traits"), fill=Traits))+
  geom_tile(aes(y=Link, x=as.factor("PhyTraits"), fill=PhyTraits))+
  geom_tile(aes(y=Link, x=as.factor("PhyLatent"), fill=PhyLatent))+
  geom_tile(aes(y=Link, x=as.factor("TraitLatent"), fill=TraitsLatent))+
  geom_tile(aes(y=Link, x=as.factor("Trio"), fill=Trio))

selection[order(-selection$sum),][,4:10]

selection$Phy <- (selection$Phy)*(1/max(selection$Phy))
selection$Latent <- (selection$Latent)*(1/max(selection$Latent))
selection$Traits <- (selection$Traits)*(1/max(selection$Traits))
selection$PhyLatent <- (selection$PhyLatent)*(1/max(selection$PhyLatent))
selection$TraitsLatent <- (selection$TraitsLatent)*(1/max(selection$TraitsLatent))
selection$PhyTraits <- (selection$PhyTraits)*(1/max(selection$PhyTraits))
selection$Trio <- (selection$Trio)*(1/max(selection$Trio))

selection <- dplyr::mutate(selection, sum=Phy+Traits+Latent+PhyTraits+TraitsLatent+PhyLatent+Trio)
selection[order(selection$sum),]

selection_0 <- dplyr::filter(selection, real==0) %>% dplyr::select(., -sum, -real, -Plant_Species, -Frugivore_Species) %>% column_to_rownames(., var="Link")

hist(selection$Trio)

heatmap(as.matrix(selection_0))
```

Variable importance

```{r}
Phy_rf <- woodedWalk_NoSplit(birds, FrugTraits = c(f_phylo), PlantTraits = c(p_phylo), class_balancing = FALSE, output_type = "rfobject")

Traits_rf <- woodedWalk_NoSplit(birds, FrugTraits = c(f_traits), PlantTraits = c(p_traits), class_balancing = FALSE, output_type = "rfobject")

Latent_rf <- woodedWalk_NoSplit(birds, FrugTraits = c(f_latent), PlantTraits = c(p_latent), class_balancing = FALSE, output_type = "rfobject")

PhyTraits_rf <- woodedWalk_NoSplit(birds, FrugTraits = c(f_phylo, f_traits), PlantTraits = c(p_phylo, p_traits), class_balancing = FALSE, output_type = "rfobject")

TraitsLatent_rf <- woodedWalk_NoSplit(birds, FrugTraits = c(f_latent, f_traits), PlantTraits = c(p_latent, p_traits), class_balancing = FALSE, output_type = "rfobject")

PhyLatent_rf <- woodedWalk_NoSplit(birds, FrugTraits = c(f_latent, f_phylo), PlantTraits = c(p_latent, p_phylo), class_balancing = FALSE, output_type = "rfobject")

Trio_rf <- woodedWalk_NoSplit(birds, FrugTraits = c(f_phylo, f_latent, f_traits), PlantTraits = c(p_phylo, p_latent, p_traits), class_balancing = FALSE, output_type = "rfobject")

models <- list("Phy_rf"=Phy_rf, "Traits_rf"=Traits_rf, "Latent_rf"=Latent_rf, "PhyTraits_rf"=PhyTraits_rf, "TraitsLatent_rf"=TraitsLatent_rf, "PhyLatent_rf"=PhyLatent_rf, 
               "Trio_rf"=Trio_rf)
```

Joint variable importance aside:
We need to one hot encode our categorical variables to keep from enforcing some sort of artifical ordination, but we don't want to interpret the importance of each dummy variable individually. If we want to find the importance of the set, we need to find the average of importance weighted  by the frequency of the variable (at least for Gini impurity according to this stack overflow post: https://stats.stackexchange.com/questions/92419/relative-importance-of-a-set-of-predictors-in-a-random-forests-classification-in)
```{r}

#' Summarize Gini impurity score across one-hot encoded variables
#' Does this by created a weighted mean impurity score, where weight is based on frequency of the variable
#'
#' @citation I DID NOT WRITE THIS CODE! Credit to Max Ghenis on Stack Overflow 
#' https://stats.stackexchange.com/questions/92419/relative-importance-of-a-set-of-predictors-in-a-random-forests-classification-in
#' 
#' @param rf.obj Random forest object you want to compute variable importance for
#' @param groups a list, where each element is a character vector of rownames you want to combine
#' 
#' @return A 1 dimensional array, where rownames are the groups of variables and associated values are weighted average importance values (accoring to Gini coefficient)
#' 
group.importance <- function(rf.obj, groups) {
var.share <- function(rf.obj, members) {
  count <- table(rf.obj$forest$bestvar)[-1]
  names(count) <- names(rf.obj$forest$ncat)
  share <- count[members] / sum(count[members])
  return(share)
}
  var.imp <- as.matrix(sapply(groups, function(g) {
    sum(randomForest::importance(rf.obj, 2)[g, ]*var.share(rf.obj, g))
  }))
  colnames(var.imp) <- "MeanDecreaseGini"
  return(var.imp)
}

group.importance(rf.obj=output[[1]]$Traits, groups=list(
                growthform=c("tree","liana", "palm", "scrub"), 
               fruitcolor=c("yellow", "red", "black", "brown", "green"),
               LipidScore=c("Lipid_Score"),
               fruit_diameter=c("fruit_diameter"),
               fruit_length=c("fruit_length"),
              Frug_Body_Mass="Frug_Body_Mass",
              Frug_Mean_Gape_Size="Frug_Mean_Gape_Size",
              FrugScore="Frugivory_score")
               )

```


```{r}

output <- data.frame(models[[7]]$importance)
output$model <- names(models)[7]
output <- rownames_to_column(output, var = "Variable")
  

for(i in 1:6){
  temp <- data.frame(models[[i]]$importance)
  temp$model <- names(models)[i]
  temp <- rownames_to_column(temp, var = "Variable")
  output <- rbind(output, temp)
}

output$MeanDecreaseGini <- as.numeric(output$MeanDecreaseGini)
```

```{r}
output$varType <- NA
output$varType[output$Variable %in% c(f_phylo, p_phylo)==TRUE] <- "Phylo"
output$varType[output$Variable %in% c(f_traits, p_traits)==TRUE] <- "Trait"
output$varType[output$Variable %in% c(f_latent, p_latent)==TRUE] <- "Latent"

output$Variable <- as.factor(output$Variable)
```

```{r}
ggplot(data=output, aes(y=Variable, x=MeanDecreaseGini, color=model))+geom_point()

plot(x=output$MeanDecreaseGini, 
     y=output$Variable, 
     yaxt = "n",
     col=as.factor(output$model), 
     pch=19)
axis(2, 1:32, levels(output$Variable))

```

```{r}


Trio_levels <- output %>% dplyr::filter(., model=="Trio_rf")
Trio_levels <- levels(reorder(Trio_levels$Variable, Trio_levels$MeanDecreaseGini))


Trio_Colors <- c("#E411E0", "#E0E411", "#11E0E4") #Phylo, Trait, Latent

colKey <- vector(mode="character", length=nrow(output))
colKey[levels(output$Variable) %in% c(f_phylo, p_phylo)==TRUE] <- Trio_Colors[1]
colKey[levels(output$Variable) %in% c(f_traits, p_traits)==TRUE] <- Trio_Colors[2]
colKey[levels(output$Variable) %in% c(f_latent, p_latent)==TRUE] <- Trio_Colors[3]

#pdf(file="VarImportance.pdf", width=11, height=11)
ggplot(data=output, aes(y=Variable, x=MeanDecreaseGini, fill=model))+geom_col(position = position_dodge(), width = 0.9)+theme_classic()+
  scale_fill_manual(values = c("Phy_rf" = "#E411E0",
                                "Traits_rf" = "#E0E411",
                                "Latent_rf" = "#11E0E4",
                                "PhyTraits_rf" = "#E27B79",
                                "TraitsLatent_rf" = "#79E27B",
                                "PhyLatent_rf" = "#7B79E2",
                                "Trio_rf" = "#56549e"
                                ))+
    theme(axis.text.y = element_text(colour = colKey, size = 15), legend.position = "none")
```

```{r}
output_niceName <- output

levels(output_niceName$Variable)[match("Psvd1",levels(output_niceName$Variable))] <- "Plant SVD: Axis 1"
levels(output_niceName$Variable)[match("Psvd2",levels(output_niceName$Variable))] <- "Plant SVD: Axis 2"
levels(output_niceName$Variable)[match("Psvd3",levels(output_niceName$Variable))] <- "Plant SVD: Axis 3"
levels(output_niceName$Variable)[match("Fsvd1",levels(output_niceName$Variable))] <- "Frugivore SVD: Axis 1"
levels(output_niceName$Variable)[match("Fsvd2",levels(output_niceName$Variable))] <- "Frugivore SVD: Axis 2"
levels(output_niceName$Variable)[match("Fsvd3",levels(output_niceName$Variable))] <- "Frugivore SVD: Axis 3"

levels(output_niceName$Variable)[match("c1PIDecomp",levels(output_niceName$Variable))] <- "Plant Phylogeny: Axis 1"
levels(output_niceName$Variable)[match("c2PIDecomp",levels(output_niceName$Variable))] <- "Plant Phylogeny: Axis 2"
levels(output_niceName$Variable)[match("c3PIDecomp",levels(output_niceName$Variable))] <- "Plant Phylogeny: Axis 3"
levels(output_niceName$Variable)[match("fc1",levels(output_niceName$Variable))] <- "Frugivore Phylogeny: Axis 1"
levels(output_niceName$Variable)[match("fc2",levels(output_niceName$Variable))] <- "Frugivore Phylogeny: Axis 2"
levels(output_niceName$Variable)[match("fc3",levels(output_niceName$Variable))] <- "Frugivore Phylogeny: Axis 3"

levels(output_niceName$Variable)[match("Frug_Mean_Gape_Size",levels(output_niceName$Variable))] <- "Frugivore Gape Size"
levels(output_niceName$Variable)[match("Frug_Body_Mass",levels(output_niceName$Variable))] <- "Frugivore Mass"
levels(output_niceName$Variable)[match("fruit_diameter",levels(output_niceName$Variable))] <- "Fruit Diameter"
levels(output_niceName$Variable)[match("fruit_length",levels(output_niceName$Variable))] <- "Fruit Length"
levels(output_niceName$Variable)[match("red",levels(output_niceName$Variable))] <- "Red Fruit"
levels(output_niceName$Variable)[match("black",levels(output_niceName$Variable))] <- "Black Fruit"
levels(output_niceName$Variable)[match("green",levels(output_niceName$Variable))] <- "Green Fruit"
levels(output_niceName$Variable)[match("yellow",levels(output_niceName$Variable))] <- "Yellow Fruit"
levels(output_niceName$Variable)[match("brown",levels(output_niceName$Variable))] <- "Brown Fruit"
levels(output_niceName$Variable)[match("tree",levels(output_niceName$Variable))] <- "Tree"
levels(output_niceName$Variable)[match("palm",levels(output_niceName$Variable))] <- "Palm"
levels(output_niceName$Variable)[match("liana",levels(output_niceName$Variable))] <- "Liana"
levels(output_niceName$Variable)[match("scrub",levels(output_niceName$Variable))] <- "Scrub"

levels(output_niceName$Variable)[match("LipScore1",levels(output_niceName$Variable))] <- "Low Lipid Content Fruit"
levels(output_niceName$Variable)[match("LipScore2",levels(output_niceName$Variable))] <- "Mid Lipid Content Fruit"
levels(output_niceName$Variable)[match("LipScore3",levels(output_niceName$Variable))] <- "High Lipid Content Fruit"

levels(output_niceName$Variable)[match("FrugScore1",levels(output_niceName$Variable))] <- "Occasional Frugivore"
levels(output_niceName$Variable)[match("FrugScore2",levels(output_niceName$Variable))] <- "Facultative Frugivore"
levels(output_niceName$Variable)[match("FrugScore3",levels(output_niceName$Variable))] <- "Strict Frugivore"

```

```{r}
#pdf(file="VarImportance.pdf", width=11, height=11)
ggplot(data=output_niceName, aes(y=Variable, x=MeanDecreaseGini, fill=model))+geom_col(position = position_dodge(), width = 0.9)+theme_classic()+
  scale_fill_manual(values = c("Phy_rf" = "#E411E0",
                                "Traits_rf" = "#E0E411",
                                "Latent_rf" = "#11E0E4",
                                "PhyTraits_rf" = "#E27B79",
                                "TraitsLatent_rf" = "#79E27B",
                                "PhyLatent_rf" = "#7B79E2",
                                "Trio_rf" = "#56549e"
                                ))+
    theme(axis.text.y = element_text(colour = colKey, size = 15), legend.position = "none")
```


```{r}
#pdf(file="VarImportanceDots.pdf", width=11, height=11)
ggplot(data=output_niceName, aes(y=Variable, x=MeanDecreaseGini, col=model))+geom_point(size=5)+theme_classic()+
  scale_color_manual(values = alpha(c("Phy_rf" = "#E411E0",
                                "Traits_rf" = "#E0E411",
                                "Latent_rf" = "#11E0E4",
                                "PhyTraits_rf" = "#E27B79",
                                "TraitsLatent_rf" = "#79E27B",
                                "PhyLatent_rf" = "#7B79E2",
                                "Trio_rf" = "#000000"
                                ), 0.5))+
    theme(axis.text.y = element_text(colour = colKey, size = 15), legend.position = "none")
```


```{r}
library(circlize)

birdsmall <- birds %>% dplyr::select(., Frug_Family, Plant_family) %>% unique()

bird_web_small <- data.frame(
    from = birdsmall$Frug_Family,
    to = birdsmall$Plant_family
  ) %>% graph_from_data_frame(directed=FALSE) #Create a graph with all unique plant-frugivore edges

adges <- get.edgelist(bird_web_small)
mat <-get.adjacency(bird_web_small, type="both", sparse = FALSE) #Create adjacency matrix from our newly created web
#isSymmetric.matrix(mat) #we should get a full symmetric adjacency matrix since we added the type="both" argument above; this makes sure

length(unique(birds$Frug_Genus))
length(unique(birds$Frugivore_Species))
length(unique(birds$Frug_Family))
#length(unique(birds$Frug_Order))

length(unique(birds$Plant_family))
length(unique(birds$Plant_family))
length(unique(birds$Plant_genus))
length(unique(birds$Plant_Species))

chordDiagram(mat)
circlize::chordDiagram(adges)
```

```{r}
library(rasterVis)
?levelplot()

levelplot(composite)
```


```{r}
# Create an adjacency matrix: 
# a list of connections between 20 origin nodes, and 5 destination nodes:
numbers <- sample(c(1:1000), 100, replace = T)
data <- matrix( numbers, ncol=5)
rownames(data) <- paste0("orig-", seq(1,20))
colnames(data) <- paste0("dest-", seq(1,5))

# Load the circlize library
library(circlize)
class(data)
# Make the circular plot
circlize::chordDiagramFromMatrix(data)
circlize::chordDiagram(data, transparency = 0.5)

```

```{r}
library(raster)
library(rasterVis)
library(scales)


composite$Frugdegree
composite$Frugivore_Species <- as.factor(composite$Frugivore_Species)
composite$Plant_Species <- as.factor(composite$Plant_Species)



composite$Frugivore_Species <- reorder(composite$Frugivore_Species, composite$Frugdegree)
composite$Plant_Species <- reorder(composite$Plant_Species, composite$Plantdegree)

range(composite$Phy/max(composite$Phy))

composite[,4:10] <- apply(composite[,4:10], 2, rescale) #use rescale to rescale all values to 0-1 for visualization


composite[composite$real==1, c("Phy", "Traits", "Latent", "PhyTraits", "TraitsLatent", "PhyLatent", "Trio")] <- NA

```

```{r}
comp_raster <- composite

comp_raster$x <- as.numeric(composite$Frugivore_Species)
comp_raster$y <- as.numeric(composite$Plant_Species)

comp_raster %<>% dplyr::select(., x, y, Phy, Traits, Latent, PhyTraits, TraitsLatent, PhyLatent, Trio)
comp_raster <- raster::rasterFromXYZ(comp_raster)
#plot(comp_raster$Trio)
#plot(comp_raster)
rasterVis::levelplot(comp_raster)



# use colorbrewer which loads with the rasterVis package to generate
# a color ramp of yellow to green

library(RColorBrewer)
cols <- colorRampPalette(brewer.pal(11,"OrRd"))
rasterVis::levelplot(comp_raster, col.regions=cols)
rasterVis::levelplot(comp_raster)

pal <- colorRampPalette(c("black", "green"))

myTheme <- BTCTheme()
myTheme$panel.background$col = 'black'

pdf(file="Draft_Figs/SuitabilityMap.pdf", width=6, height=6)
levelplot(comp_raster, col.regions=c("white", rev(viridisLite::inferno(15)), '#ffb6c1'), par.settings = myTheme, layout=c(3, 2), scales=list(tck=c(0,0), draw=FALSE))
dev.off()


pdf(file="Draft_Figs/SuitabilityMap_Trio.pdf", width=6, height=6)
rasterVis::levelplot(comp_raster$Trio, col.regions=c("white", rev(viridis::inferno(15)), '#ffb6c1'), par.settings = myTheme, scales=list(tck=c(0,0), draw=FALSE), colorkey=FALSE)
dev.off()

```

