---
title: "Plug&Play Troubleshooting"
author: "Grant Foster"
date: "10/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Functionalizing Plug n Play
```{r}
pp.kde <- function(dat, FrugTraits, PlantTraits){
  #Set up our data into a fully expanded edgelist
  require(tidyr)
  dat$real <- 1 #make a new column denoting all of these edges are real; important when we expand out to all pairwise connections later
  dat <- dplyr::filter_at(dat, vars(c(FrugTraits, PlantTraits)), all_vars(!is.na(.))) #make sure we have data for all our predictors
  full_L <- tidyr::expand(dat, Frugivore_Species, Plant_Species) #Expand to include all possible pairwise plant-frugivore interactions.
  full_real <- dplyr::select(dat, Frugivore_Species, Plant_Species, real) %>% left_join(full_L, ., by = c("Frugivore_Species", "Plant_Species")) #notating which of our edges in this expanded data are real 
  full_real$real[is.na(full_real$real)==TRUE] <- 0

  full_real_frugs <- dat[,c("Frugivore_Species", FrugTraits)] %>% unique() %>% left_join(full_real, ., by="Frugivore_Species") #Add in our frugivore traits

  full_real_both <- dat[,c("Plant_Species", PlantTraits)] %>% unique() %>% left_join(full_real_frugs, ., by="Plant_Species") #add in our plant traits
  
  # Subsample our edgelist
  sample_IDs <- sample(round(nrow(full_real_both)*0.8), replace=F) #subsample out 80% of our edges. 
  test <- full_real_both[-sample_IDs,] #assign 80% train set
  train <- full_real_both[sample_IDs,]

  #set up our data to go into npudens
  p <- train[(train[,'real'] == 1), 4:ncol(train)] #positives cases (p)
  bgrd <- train[,4:ncol(train)] #background (positive and negative) cases

  require(np)
  f1 <- np::npudens(p, bwmethod='normal-reference') 
  f0 <- np::npudens(bgrd, bwmethod='normal-reference')
  model=list(f1, f0, test, train)
  names(model) <- c("f1", "f0", "test", "train")
  #browser
  return(model)
}


predict.pp.kde <- function(model, x){
  require(np)
  f1 <- fitted(npudens(bws = model$f1$bws, edat = x, bwmethod = 'normal-reference'))
  f0 <- fitted(npudens(bws = model$f0$bws, edat = x, bwmethod = 'normal-reference'))
  #browser()
  return(S = f1/f0)
}



bats_kde <- pp.kde(bats, FrugTraits = c("Frug_Body_Mass","Frug_Mean_Gape_Size", "FrugSc1", "FrugSc2", "FrugSc3"), PlantTraits = c("fruit_diameter", "fruit_length", "tree", "palm", "scrub", "yellow", "red", "black", "brown", "green", "LipSc1", "LipSc2"))

bats_pred <- predict.pp.kde(model=bats_kde, x=bats_kde$test[,4:ncol(bats_kde$test)])
bats_kde$test$S <- bats_pred

test_results <- bats_kde$test
table(is.nan(test_results$S))
#out %<>% filter(., is.nan(S)==FALSE)

#out$S[is.infinite(out$S)] <- max(out$S[is.infinite(out$S)==FALSE], na.rm = TRUE)+1
AUC_test <- roc(data=test_results, response=real, predictor=S)
plot(AUC_test)
AUC_test
```



```{r}

birds <- dplyr::filter(dat, Frug_Class=="Aves")

birds_kde <- pp.kde(birds, FrugTraits = c("Frug_Body_Mass","Frug_Mean_Gape_Size"), PlantTraits = c("fruit_diameter", "fruit_length", "tree", "palm", "scrub", "yellow", "red", "black", "brown", "green", "LipSc1", "LipSc2"))

birds_pred <- predict.pp.kde(model=birds_kde, x=birds_kde$test[,4:ncol(birds_kde$test)])
birds_kde$test$S <- birds_pred

test_results <- birds_kde$test
table(is.nan(test_results$S))
#out %<>% filter(., is.nan(S)==FALSE)

#out$S[is.infinite(out$S)] <- max(out$S[is.infinite(out$S)==FALSE], na.rm = TRUE)+1
AUC_test <- roc(data=test_results, response=real, predictor=S)
plot(AUC_test)
AUC_test
```


### Below here is scratch work on Odd path counting; I'll likely delete later!
Creating link prediction function

Note: If I make a strictly upper triangular matrix (non-symmetric), it's eigenvalues will all be 0. So we need to see a symmetric matrix. 

We want ideally to minimize the Sigma[f(delta_kk)-U^T * A' * Uk]^2, essentially a sum of squares.
```{r}
U <-eigen(mat)$vectors
L <- eigen(mat)$values
table(mat)
table(mat)/sum(table(mat))


# A = U*L*U^T
test <- U %*% diag(L) %*% t(U)
test <- round(test, digits=3) #Rounding errors make our 0's into super small values
table(test==mat) #This should be true

#A' = F(A) = U*F(L)*U^T, where F is our kernel function relating our training set (A) to test set (A')
#We want to minimize |F(A)-A'|, where

U_train <- eigen(test_adj)$vectors
L_train <- eigen(test_adj)$values

target_V <- t(U_train) %*% mat %*% U_train
target_V <- diag(target_V)

plot(x=L_train, y=target_V, col="dodger blue", pch=16)
```


The first kernel that we use is odd path counting. Here we look at weighted sum of several odd powers of the original adjacency matrix (recall that k-th power of an adjacency matrix refers to numbers of k-hop neighbours from every node). The higher the value, the more likely it is for two vertices to be connected.

I think this is mostly wrong. I hope to fix it in the future.

```{r}
library(expm)

odd_kernal <- function(input, max_path){
  index <- seq(from=1, to=max_path, by=2)
  #index <- index[-1]
  output <- 0
  for(i in index){
    #print(i)
    output <- output + (test_adj %^% i)
  }
  return(output)
}

odd_paths_V1 <- t(U_train) %*% odd_kernal(input=test_adj, max_path = 1) %*% U_train
odd_paths_pred1 <- diag(odd_paths_V1)

odd_paths_V3 <- t(U_train) %*% odd_kernal(input=test_adj, max_path = 3) %*% U_train
odd_paths_pred3 <- diag(odd_paths_V3)

odd_paths_V5 <- t(U_train) %*% odd_kernal(input=test_adj, max_path = 5) %*% U_train
odd_paths_pred5 <- diag(odd_paths_V5)

odd_paths_V7 <- t(U_train) %*% odd_kernal(input=test_adj, max_path = 7) %*% U_train
odd_paths_pred7 <- diag(odd_paths_V7)

odd_paths_V9 <- t(U_train) %*% odd_kernal(input=test_adj, max_path = 9) %*% U_train
odd_paths_pred9 <- diag(odd_paths_V9)

plot(x=L_train, y=target_V, col="dodger blue", pch=16)
points(x=L_train, y=odd_paths_pred1, col="red", pch=16)
points(x=L_train, y=odd_paths_pred3, col="orange", pch=16)
points(x=L_train, y=odd_paths_pred5, col="green", pch=16)
points(x=L_train, y=odd_paths_pred7, col="purple", pch=16)
points(x=L_train, y=odd_paths_pred9, col="navy", pch=16)
legend(1, 1, legend=c("1 (base)", "3", "5", "7", "9"),
       col=c("red", "orange", "green", "purple", "navy"), pch=16, cex=0.8)

all.equal(dim(odd_paths_V1), dim(mat))

edgelist <- data.frame(matrix(mat, dimnames=list(t(outer(colnames(mat), rownames(mat), FUN=paste)), NULL)))
colnames(edgelist) <- "truth"
edgelist <- rownames_to_column(edgelist, var="link")


colnames(odd_paths_V1) <- colnames(mat)
rownames(odd_paths_V1) <- rownames(mat)
odd_paths_1L <- data.frame(matrix(odd_paths_V1, dimnames=list(t(outer(colnames(odd_paths_V1), rownames(odd_paths_V1), FUN=paste)), NULL)))
colnames(odd_paths_1L) <- "Path1"
odd_paths_1L <- rownames_to_column(odd_paths_1L, var="link")

colnames(odd_paths_V3) <- colnames(mat)
rownames(odd_paths_V3) <- rownames(mat)
odd_paths_3L <- data.frame(matrix(odd_paths_V3, dimnames=list(t(outer(colnames(odd_paths_V3), rownames(odd_paths_V3), FUN=paste)), NULL)))
colnames(odd_paths_3L) <- "Path3"
odd_paths_3L <- rownames_to_column(odd_paths_3L, var="link")

ROC_data <- left_join(edgelist, odd_paths_1L, by="link")
ROC_data <- left_join(ROC_data, odd_paths_3L, by="link")

simple_roc <- function(labels, scores){
  labels <- labels[order(scores, decreasing=TRUE)]
  data.frame(TPR=cumsum(labels)/sum(labels), FPR=cumsum(!labels)/sum(!labels), labels)
}

temp <- simple_roc(labels=ROC_data$truth, scores=ROC_data$Path1)
temp2 <- simple_roc(labels=ROC_data$truth, scores=ROC_data$Path3)
plot(x=temp$FPR, y=temp$TPR, type="l", col="blue")
lines(x=temp2$FPR, y=temp2$TPR, type="l", col="red")


AUC_Odd1 <- roc(response=ROC_data$truth, predictor=ROC_data$Path1)
AUC_Odd3 <- roc(response=ROC_data$truth, predictor=ROC_data$Path3)

auc(AUC_Odd1)
plot(AUC_Odd1)

auc(AUC_Odd3)
plot(AUC_Odd3)
```

  
  

Beta*sinh(alpha*A)


alpha * (np.exp(lamb) - np.exp(-lamb)) for lamb in V
x=v_train
y=target_V

y = function(x) + eps
V_train




```{r}
rgraph <-bipartite.random.game(15, 10, type="gnm", m=50)

rgraph$type <- layout_as_bipartite(rgraph)
plot(rgraph, layout=layout_as_bipartite(rgraph))

train_Edges <- sample(1:length(E(rgraph)), round(length(E(rgraph))*0.7), replace = FALSE)


test <- igraph::delete.edges(rgraph, train_Edges)


x <- x[x %in% train_Edges ==FALSE]
train <- igraph::delete.edges(rgraph, x)

plot(train, layout=layout_as_bipartite(train))
plot(test, layout=layout_as_bipartite(test))

```


```{r}
alphas <- seq(from=-1, to=1, by=0.1)

result <- NULL
for(i in 1:length(alphas)){
temp <- alphas[i]*(expm::expm(diag(L))) - expm::expm(diag(L))
result[i] <- sum((temp-U_train)^2)
}

alphas[which.min(result)]
result

test_A <- get.adjacency(test, sparse=FALSE)
train_A <- get.adjacency(train, sparse=FALSE)
?get.adjacency
z <- eigen(train_A)

target <- vector()
for(i in 1:ncol(test_A)){
target[i] <- t(eigen(train_A)$vector)[i,] %*% test_A %*% eigen(train_A)$vector[,i] #UkTA'Uk
}


plot(x=eigen(train_A)$values, y=target)
```

